---
title: "APAL_data_analysis"
output: html_document
date: "2025-10-31"
---

README: analyzing the standardized A. palmata data to estimate RSE model parameters

```{r setup, include=FALSE}

# command + option + i for new code chunk
# command + shift + enter to run entire code chunk
# command + enter to run single line of code

library(tidyverse) # command + shift + m for %>%

library(readxl)

library(janitor)

library("maps") # https://rdrr.io/cran/maps/man/

library("sf")

library("ggspatial")# for adding scale bar and north arrow to maps


```


to get variability from proportions, just take random subsets of the data, calculate proportions surviving, and repeat? compare to binomial GLMM

histograms or ridgeline plots (separate for each size class?) where x = proportion surviving and proportions are calculated by grouping values by study, region, fragment or not, etc?

# import data

```{r}

apal_surv_ind <- read.csv("../../standardized_data/apal_surv_ind.csv") # individual survival
apal_growth_ind <- read.csv("../../standardized_data/apal_growth_ind.csv") # individual growth rates

apal_surv_summ <- read.csv("../../standardized_data/apal_surv_summ.csv") # survival from studies that didn't report individual data
apal_growth_summ <- read.csv("../../standardized_data/apal_growth_summ.csv") # growth rates from studies that didn't report individual data


# update sizes for the individual data (use live cm2 where available; if unavailable then just use skeletal cm2 and assume 100% live cover)
apal_surv_ind <- apal_surv_ind %>% mutate(size_cm2_1  = if_else(is.na(size_live_cm2)==F, size_live_cm2, size_cm2))
apal_growth_ind <- apal_growth_ind %>% mutate(size_cm2_1  = if_else(is.na(size_live_cm2)==F, size_live_cm2, size_cm2))

```


# define size classes

```{r}

# lower bounds of each size class (cm2)
SC1 <- 0
SC2 <- 10
SC3 <- 100
SC4 <- 900
SC5 <- 4000

SC_lower <- c(SC1, SC2, SC3, SC4, SC5)
SC_upper <- c(SC2, SC3, SC4, SC5, max(c(apal_surv_ind$size_live_cm2, apal_surv_summ$size_cm2_max), na.rm = T))

#quantile(c(apal_surv_ind$size_live_cm2[which(apal_surv_ind$size_live_cm2 > SC5)], apal_surv_summ$size_cm2_max[which(apal_surv_summ$size_cm2_max > SC5)]), 0.5)

# function for assigning sizes to size classes
SC_fun <- function(size){
  
  if(is.na(size)==F){
    if(size <= SC2){
      size_class <- 1
    } 
    
    if(size <= SC3 & size > SC2){
      size_class <- 2
    }
    
    if(size <= SC4 & size > SC3){
      size_class <- 3
    }
    
    if(size <= SC5 & size > SC4){
      size_class <- 4
    }
    
    if(size > SC5){
      size_class <- 5
    }
  } else{
    size_class <- NA
  }
  
  return(size_class)
}

# vectorize this function
SC_funV <- Vectorize(SC_fun, "size")

```

# prep data



Dealing with non-annual time intervals in the individual survival data
-- since we are interested in ~annual survival rates, need to deal with cases in the individual survival data where the sample interval was <11 months or >13 months  

-- if sampling interval was less than a year and the coral died, know that survival at exactly one year would also have been 0
-- if sampling interval was greater than a year and the coral survived, know that survival at exactly 1 year would have also been 1 
-- for the other cases (survived less than a year or died after more than a year) use proportions surviving from that region/year/size class to estimate survival rates over the sampling interval, then convert to annual survival rates and move these to the summary data set


```{r}

#apal_surv_ind %>% filter(time_interval_yr < 11/12) %>% view()
#apal_surv_ind %>% filter(time_interval_yr > 13/12) %>% view()

apal_surv_ind1 <- apal_surv_ind %>% mutate(time_interval_yr = if_else(time_interval_yr < 11/12 & survived==0, 1, if_else(time_interval_yr > 13/12 & survived==1, 1, time_interval_yr))) 

# add a column with colony size class to the dataset
apal_surv_ind1$size_class <- rep(NA, nrow(apal_surv_ind1))

for(i in 1:nrow(apal_surv_ind1)){
  
  if(is.na(apal_surv_ind1$size_live_cm2[i])==F){
    apal_surv_ind1$size_class[i] <- SC_fun(apal_surv_ind1$size_live_cm2[i])
  }
  
  if(is.na(apal_surv_ind1$size_live_cm2[i])==T & is.na(apal_surv_ind1$size_cm2[i])==F){
    apal_surv_ind1$size_class[i] <- SC_fun(apal_surv_ind1$size_cm2[i])
  }
  
}


#View(apal_surv_ind1)

apal_surv_ind1_summ <- apal_surv_ind1 %>% group_by(study, region, survey_yr, size_class, data_type) %>% summarise(n_total = n(), n_survived = sum(survived == 1)) %>% mutate(prop_survived = n_survived/n_total)

#View(apal_surv_ind1_summ)

# observations where time interval is less than ~1 yr and coral was alive at time of sampling
under_yr <- apal_surv_ind1 %>% filter(time_interval_yr < 11/12) %>% group_by(study, region, survey_yr, data_type) %>% mutate(time_interval_yr_mean = mean(time_interval_yr), time_interval_yr_sd = sd(time_interval_yr)) %>% ungroup() %>% group_by(study, region, survey_yr, size_class, data_type, time_interval_yr_mean, time_interval_yr_sd) %>% summarize(n_unknown = n())

#View(under_yr)

# observations where time interval is greater than ~1 yr and coral was dead at time of sampling
over_yr <- apal_surv_ind1 %>% filter(time_interval_yr > 13/12) %>% group_by(study, region, survey_yr, data_type) %>% mutate(time_interval_yr_mean = mean(time_interval_yr), time_interval_yr_sd = sd(time_interval_yr)) %>% ungroup() %>% group_by(study, region, survey_yr, size_class, data_type, time_interval_yr_mean, time_interval_yr_sd) %>% summarize(n_unknown = n())

#View(over_yr)

under_yr_summ <- under_yr %>% left_join(apal_surv_ind1_summ, by = c("study", "region", "survey_yr", "size_class", "data_type")) %>% rename(n_initial = n_unknown, n_reference = n_total) %>% select(-n_survived) %>% mutate(prop_survived = prop_survived^(1/time_interval_yr_mean))

over_yr_summ <- over_yr %>% left_join(apal_surv_ind1_summ, by = c("study", "region", "survey_yr", "size_class", "data_type")) %>% rename(n_initial = n_unknown, n_reference = n_total) %>% select(-n_survived) %>% mutate(prop_survived = prop_survived^(1/time_interval_yr_mean))


#View(under_yr_summ)
#View(over_yr_summ)

# join these together
apal_surv_ind_summ <- rbind(under_yr_summ, over_yr_summ)

#View(apal_surv_ind_summ)

# now add the individual sizes/metadata for all of the colonies with unknown fates
apal_ind_summ_sizes <- apal_surv_ind1 %>% filter(time_interval_yr < 11/12 | time_interval_yr > 13/12) %>% select(-survived)#select(study, region, survey_yr, size_class, size_live_cm2)

#View(apal_ind_summ_sizes)

apal_surv_ind_summ <- apal_surv_ind_summ %>% left_join(apal_ind_summ_sizes, by = c("study", "region", "survey_yr", "size_class", "data_type"), relationship = "many-to-many")

#View(apal_surv_ind_summ)


# now remove the colonies with unknown fates from the individual survival data
apal_surv_ind1 <- apal_surv_ind1 %>% filter(time_interval_yr >= 11/12 & time_interval_yr <= 13/12)

```


address limited/missing information in the summarized survival data:

```{r}

#View(apal_surv_summ)

# exclude data sets where only a range of sizes is given (no information about the mean) and that range spans multiple defined size classes


apal_surv_summ1 <- apal_surv_summ

apal_surv_summ1$size_class_min <- rep(NA, nrow(apal_surv_summ1))
apal_surv_summ1$size_class_max <- rep(NA, nrow(apal_surv_summ1))
apal_surv_summ1$size_class_mean <- rep(NA, nrow(apal_surv_summ1))

for(i in 1:nrow(apal_surv_summ1)){
    
    # size class of min size
  apal_surv_summ1$size_class_min[i] <- SC_fun(apal_surv_summ1$size_cm2_min[i])
  
    # size class of max size
  apal_surv_summ1$size_class_max[i] <- SC_fun(apal_surv_summ1$size_cm2_max[i])
    
    # size class of mean size
  apal_surv_summ1$size_class_mean[i] <- SC_fun(apal_surv_summ1$size_cm2_mean[i])
  
}

#View(apal_surv_summ1)

# now filter out cases where there is no mean and the min and max sizes are in different size classes
apal_surv_summ1 <- apal_surv_summ1 %>% mutate(size_flag = if_else(is.na(size_cm2_mean)==T & size_class_min!=size_class_max, 1, 0)) %>% filter(size_flag == 0) %>% select(-size_flag)


#View(apal_surv_summ1)

```



address limited/missing information in the growth data

```{r}

# for the summarized data, remove cases where there is no mean and the min and max sizes are in different size classes
apal_growth_summ1 <- apal_growth_summ %>% mutate(size_class_min = SC_funV(size_cm2_min), size_class_max = SC_funV(size_cm2_max), size_class_mean = SC_funV(size_cm2_mean)) %>% mutate(size_flag = if_else(is.na(size_cm2_mean)==T & size_class_min!=size_class_max, 1, 0)) %>% filter(size_flag == 0) %>% select(-size_flag) 

#View(apal_growth_summ1)

# update sizes for the individual data (use live cm2 where available; if unavailable then just use skeletal cm2 and assume 100% live cover)
apal_growth_ind2 <- apal_growth_ind %>% mutate(size_cm2_1  = if_else(is.na(size_live_cm2)==F, size_live_cm2, size_cm2)) %>% mutate(growth_cm2_yr_1 = if_else(is.na(growth_live_cm2_yr)==F, growth_live_cm2_yr, growth_cm2_yr))

```


## data summary

number of studies, regions, etc.

```{r}

# survival
length(unique(c(apal_surv_ind$study, apal_surv_summ$study)))
length(unique(c(apal_surv_ind$region, apal_surv_summ$region)))
unique(c(apal_surv_ind$region, apal_surv_summ$region))

# growth
length(unique(c(apal_growth_ind$study, apal_growth_summ$study)))
length(unique(c(apal_growth_ind$region, apal_growth_summ$region)))
unique(c(apal_growth_ind$region, apal_growth_summ$region))

```



## map


```{r}

world <- map_data('world')

latlon_survI <- apal_surv_ind %>% group_by(study, latitude, longitude) %>% summarize()
latlon_growthI <- apal_growth_ind %>% group_by(study, latitude, longitude) %>% summarize()
latlon_survS <- apal_surv_summ %>% filter(is.na(latitude)==F) %>% group_by(study, latitude, longitude) %>% summarize()
latlon_growthS <- apal_growth_summ %>% filter(is.na(latitude)==F) %>% group_by(study, latitude, longitude) %>% summarize()

#View(latlon_growthI)

study_latlon <- rbind(latlon_survI, latlon_growthI, latlon_survS, latlon_growthS) %>% distinct() %>% filter(study != "williams_miller_2010") # remove duplicate rows and Williams & Miller 2010 (which doesn't get used later because the size intervals are too big)

# get study, lat, and lon for each df, then rbind and remove duplicates

study_latlon_sf <- study_latlon %>% st_as_sf(coords = c("longitude","latitude")) %>% st_set_crs(4326)

#View(study_latlon_sf)

```


```{r}

#max(surv.latlon$long)
#max(surv.latlon$lat)

# https://sronpersonalpages.nl/~pault/

s_cols <- c("#D9CCE3", "#CAACCB", "#BA8DB4", "#AA6F9E", "#994F88", "#882E72", "#1965B0", "#437DBF", "#6195CF", "#7BAFDE", "#4EB265", "#90C987", "#CAE0AB", "#F7F057", "#F7CB45", "#F4A736", "#EE8026", "#E65518", "#DC050C", "#A5170E", "#72190E")

study_list <- unique(c(apal_surv_ind1$study, apal_surv_summ1$study, apal_growth_ind$study, apal_growth_summ$study)) # all the studies in apal_surv_ind_summ are in apal_surv_ind2

col_df <- data.frame(
  study = sort(study_list), # sort alphabetically?
  cols = s_cols[1:length(study_list)]
  
)


#View(col_df)

map.colors <- c(bruckner_bruckner_2001 = s_cols[1], chamberland_et_al_2015= s_cols[2], forrester_et_al_2013= s_cols[3], fundemar_fragments= s_cols[4], fundemar_recruits= s_cols[5], garrison_ward_2008= s_cols[6], kuffner_et_al_2020= s_cols[7], maurer_et_al_2022= s_cols[8], mendoza_quiroz_et_al_2023= s_cols[9], NOAA_survey= s_cols[10], ortiz_prosper_2005= s_cols[11], papke_et_al_2021= s_cols[12], pausch_et_al_2018= s_cols[13], rosales_et_al_2024= s_cols[14], roth_et_al_2013= s_cols[15],schutter_et_al_2023 = s_cols[16], USGS_USVI_exp= s_cols[17], vardi_2011= s_cols[18]) #, williams_miller_2010 = NA

#png("figures/study_map.png")
study_map <- ggplot() + 
  geom_polygon(data = world, aes(x=long, y = lat, group = group),colour='black',fill='gray70', linewidth = 0.2) + 
  geom_sf(data = study_latlon_sf, aes(fill = study), color = "black", size = 3, shape=21)+
  coord_sf(xlim = c(-90, -62), ylim = c(10, 27)) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_rect(fill = 'gray95', color = 'gray95'), axis.text.x = element_text(size = 12), axis.text.y = element_text(size = 12), axis.title = element_blank(), panel.border = element_rect(colour = "black", fill=NA, linewidth=0.8), plot.title = element_text(hjust=0, size = 14))+
  scale_fill_manual(values=map.colors)+
  #guides(fill=guide_legend(ncol=2)) +
  annotation_scale(location = "tr", style = "ticks",  pad_x = unit(0.25, "cm"), pad_y = unit(0.5, "cm"), text_cex=1)+#text_cex=0.9, pad_x = unit(0.25, "cm"), pad_y = unit(0.45, "cm")
  annotation_north_arrow(location = "tl", style = north_arrow_orienteering(text_size = 11), height = unit(0.6, "cm"), width = unit(0.6, "cm"), pad_y = unit(0.05, "cm"))
#dev.off()

ggsave("figures/study_map.png", width = 7, height = 5)


```

note the point between Haiti and Jamaica is Navassa Island


update: color by region

```{r}


latlon_survI <- apal_surv_ind %>% group_by(study, latitude, longitude, region) %>% summarize()
latlon_growthI <- apal_growth_ind %>% group_by(study, latitude, longitude, region) %>% summarize()
latlon_survS <- apal_surv_summ %>% filter(is.na(latitude)==F) %>% group_by(study, latitude, longitude, region) %>% summarize()
latlon_growthS <- apal_growth_summ %>% filter(is.na(latitude)==F) %>% group_by(study, latitude, longitude, region) %>% summarize()

#View(latlon_growthI)

study_latlon2 <- rbind(latlon_survI, latlon_growthI, latlon_survS, latlon_growthS) %>% distinct() %>% filter(study != "williams_miller_2010") # remove duplicate rows and Williams & Miller 2010 (which doesn't get used later because the size intervals are too big)

#View(study_latlon2)

# get study, lat, and lon for each df, then rbind and remove duplicates

study_latlon_sf2 <- study_latlon2 %>% st_as_sf(coords = c("longitude","latitude")) %>% st_set_crs(4326)

#View(study_latlon_sf2)

m_cols <- c("#D9CCE3",  "#BA8DB4", "#882E72", "#1965B0", "#7BAFDE", "#4EB265", "#CAE0AB", "#F7F057", "#F4A736",  "#E65518", "#DC050C", "#A5170E")

region_list <- unique(c(apal_surv_ind1$region, apal_surv_summ1$region, apal_growth_ind$region, apal_growth_summ$region)) # all the studies in apal_surv_ind_summ are in apal_surv_ind2

Rcol_df <- data.frame(
  region = sort(region_list), # sort alphabetically?
  cols = m_cols[1:length(region_list)]
  
)


#View(col_df)

map.colors <- c("Bahamas" = m_cols[1], "British Virgin Islands"= m_cols[2], "Curacao"= m_cols[3], "Dominican Republic"= m_cols[4], "Dry Tortugas"= m_cols[5], "Florida Keys"= m_cols[6], "Jamaica"= m_cols[7], "Mexican Caribbean"= m_cols[8], "Navassa"= m_cols[9], "Puerto Rico"= m_cols[10], "USVI"= m_cols[11], "Virgin Gorda"= m_cols[12]) 

#png("figures/study_map.png")
study_map <- ggplot() + 
  geom_polygon(data = world, aes(x=long, y = lat, group = group),colour='black',fill='gray70', linewidth = 0.2) + 
  geom_sf(data = study_latlon_sf2, aes(fill = region), color = "black", size = 3, shape=21)+
  coord_sf(xlim = c(-90, -62), ylim = c(10, 27)) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_rect(fill = 'gray95', color = 'gray95'), axis.text.x = element_text(size = 12), axis.text.y = element_text(size = 12), axis.title = element_blank(), panel.border = element_rect(colour = "black", fill=NA, linewidth=0.8), plot.title = element_text(hjust=0, size = 14))+
  scale_fill_manual(values=map.colors)+
  #guides(fill=guide_legend(ncol=2)) +
  annotation_scale(location = "tr", style = "ticks",  pad_x = unit(0.25, "cm"), pad_y = unit(0.5, "cm"), text_cex=1)+#text_cex=0.9, pad_x = unit(0.25, "cm"), pad_y = unit(0.45, "cm")
  annotation_north_arrow(location = "tl", style = north_arrow_orienteering(text_size = 11), height = unit(0.6, "cm"), width = unit(0.6, "cm"), pad_y = unit(0.05, "cm"))
#dev.off()

ggsave("figures/region_map.png", width = 7, height = 5)


# number of studies in each region
study_latlon3 <- study_latlon2 %>% ungroup() %>% select(study, region) %>% distinct() %>% group_by(region) %>% summarize(n = n())

study_latlon3

```


# data analysis


## survival

make a function to turn summarized/proportional survival data into a data frame with randomly drawn individual sizes and survival


```{r}

# test using a while loop to draw values within a specified range from a normal distribution

# x <- which(10 == 0)
# 
# while(length(x) < 10){
#   
#   x <- rnorm(10, 10, 5)
#   
#   x <- x[which(x>0 & x < 15)]
#   
# }
# 
# x

```


```{r}

# prop_surv_df = data frame with survival expressed as proportion of individuals alive at the end of the time point
# n_reps = number of realizations (data sets) to make
# seed_n = seed for random number generation

survi_fun <- function(prop_surv_df, n_reps, seed_n){
  
  set.seed(seed_n) # set the seed
  
  data_list <- list() # holding list for the data sets
  
  for(i in 1:n_reps){ # for each realization
    
   for(j in 1:nrow(prop_surv_df)){
     
     # number of individuals that went into the jth observation of proportion surviving
     n_ind <- prop_surv_df$n_initial[j]
     
     # make the holding data set to fill in with individual values (and keep all the metadata)
     data.ij <- prop_surv_df %>% select(study, region, location, plot, treatment_1, treatment_2, latitude, longitude, depth_m, survey_yr, data_type, size_live, fragment, time_interval_yr, disturbance)
     data.ij <- data.ij[j,] # want just the jth row
     
     # replicate all the metadata n_ind times
     data.ij <- data.ij[rep(seq_len(nrow(data.ij)), n_ind), ]
     
     # now fill in size and individual survival columns
     data.ij$survived <- rbinom(n_ind, 1, prop_surv_df$prop_survived[j])
     
     # holding column for size:
     data.ij$size_cm2 <- rep(NA, nrow(data.ij))
     
     # fill in sizes by drawing a random size for each individual
      # if there is no information on variability (size_sd = NA)
     if(is.na(prop_surv_df$size_cm2_sd[j])==T){
       
       # if only mean size is known, just draw individual survival rates and assume they all have the same size
       if(is.na(prop_surv_df$size_cm2_min[j])==T){
         data.ij$size_cm2 <- rep(prop_surv_df$size_cm2_mean[j], n_ind)
         
       } else{ # if only range is known or if only mean/midpoint and range is known, draw from uniform distribution
         data.ij$size_cm2 <- runif(n_ind, min = prop_surv_df$size_cm2_min[j], max = prop_surv_df$size_cm2_max[j])
       }
      
       
     } else{ # if there is information on size variability
       
       if(is.na(prop_surv_df$size_cm2_min[j])==F){ # if mean size, sd, and range are known, draw sizes from a normal distribution and adjust to make sure they are within the given range
         
         x1 <- which(10 == 1) # start with length 0
         
         while(length(x1) < n_ind){
           # draw sizes from a normal distribution
           x1 <- c(x1, rnorm(n_ind-length(x1), prop_surv_df$size_cm2_mean[j], prop_surv_df$size_cm2_sd[j]))
           # make sure these values all fall within the observed size range:
           x1 <- x1[which(x1>= prop_surv_df$size_cm2_min[j] & x1 <= prop_surv_df$size_cm2_max[j])]
         }
         
         data.ij$size_cm2 <- x1
         
       } else{ # if mean size and sd are known but there is no information on range, draw sizes from a normal distribution and adjust to make sure they are non-zero
         
         x1 <- which(10 == 1) # start with length 0
         
         while(length(x1) < n_ind){
           # draw sizes from a normal distribution
           x1 <- c(x1, rnorm(n_ind-length(x1), prop_surv_df$size_cm2_mean[j], prop_surv_df$size_cm2_sd[j]))
           # make sure these values all fall within the observed size range:
           x1 <- x1[which(x1>= 0)]
         }
         
         data.ij$size_cm2 <- x1
         
       }
     
       
     }
     
     # now add a column with size class to the dataset
     data.ij$size_class <- SC_funV(data.ij$size_cm2)
     
     if(j == 1){
       data.j <- data.ij
     } else{
       
       data.j <- rbind(data.j, data.ij)
       
     }
    
     
   } # end of iteration over rows in the data frame
    
    # store this data frame in the list
    data_list[[i]] <- data.j
    
  } # end of iteration over random replicates
  
  
  # return the list of data frames
  
  return(data_list)
}


# for the summary survival data for the colonies with unknown annual fates from the individual survival data, know each individual colony's size so it's just a coin flip of whether that individual survived

surv_indi_fun <- function(surv_ind_summ_df, n_reps, seed_n){

  set.seed(seed_n) # set the seed
  
  data_list <- list() # holding list for the data sets
  
  
  for(i in 1:n_reps){
    
    data.i <- surv_ind_summ_df
    data.i$survived <- rep(NA, nrow(data.i))
    
    for(j in 1:nrow(data.i)){
      data.i$survived[j] <- rbinom(1, 1, data.i$prop_survived[j])
    }
    
    data_list[[i]] <- data.i
     
  }

  return(data_list)

}



```


test the functions

```{r}

test1 <- survi_fun(apal_surv_summ1, n_reps = 1, seed_n = 100)

# nrow(test1[[1]])
# 
#View(test1[[1]])

test2 <- surv_indi_fun(apal_surv_ind_summ, n_reps = 1, seed_n = 200)
#View(test2[[1]])

```




plot the data as 0s/1s

```{r}

# need to filter out only the data points with ~annual time intervals, and need to make a column with size to use (size_live_cm2 if available, otherwise size_cm2)


# open points: survival and/or size were randomly drawn based on available information on proportions surviving and size ranges
f.pch <- 16 # field data
n.pch <- 23 # nursery data
l.pch <- 17 # lab data

f.pchR <- 1 # field data, random
n.pchR <- 5 # nursery data, random
l.pchR <- 2 # lab data, random

alpha1 <- 0.25

jt_sd <- 0.01 # how much to jitter the points

pch_fun <- function(data_type, study_type){
  
  if(study_type == "I"){
    pch_out <- ifelse(data_type == "field", f.pch, ifelse(data_type == "nursery", n.pch, l.pch))
  }
  
  if(study_type == "S"){
    pch_out <- ifelse(data_type == "field", f.pchR, ifelse(data_type == "nursery", n.pchR, l.pchR))
  }
  
  return(pch_out)
}

plot_dfI <- apal_surv_ind1 %>% rename(size_cm2_p = size_cm2_1) %>% filter(is.na(disturbance)==T) # individual data
plot_dfS <- test1[[1]] %>% rename(size_cm2_p = size_cm2)%>% filter(is.na(disturbance)==T)  # individual survival and size values drawn from summarized data
plot_dfIS <- test2[[1]] %>% rename(size_cm2_p = size_live_cm2)%>% filter(is.na(disturbance)==T)  # individual annual survival data drawn from summarized individual data

plot(x = plot_dfI$size_cm2_p, y = plot_dfI$survived, type = "p", pch = 16, col = NA, las = 1, xlab = NA, ylab = NA)
mtext(side = 1, "Size (cm^2)", line = 2.5)
mtext(side = 2, "Prop. surviving", line = 2.5)
axis(side = 1, at = c(SC1, SC2, SC3, SC4, SC5), labels = NA, lwd.ticks = 2, col.ticks = "red")
# studies with individual data points
study_list1 <- study_list[which(study_list %in% unique(plot_dfI$study))]
for(i in 1:length(study_list1)){ # for each study
  study.i <- plot_dfI$study[which(plot_dfI$study==study_list1[i])][1]
  types.i <- unique(plot_dfI$data_type[which(plot_dfI$study==study.i)])
  for(j in 1:length(types.i)){ # for each data type
    types.ij <- types.i[j]
    dt_sub <- plot_dfI[which(plot_dfI$study==study.i & plot_dfI$data_type==types.i[j]),]
    jt <- rnorm(nrow(dt_sub), 0, jt_sd) # jitter the points
    lines(x = dt_sub$size_cm2_p, y = dt_sub$survived + jt, type = "p", pch = pch_fun(types.ij, "I"), col = adjustcolor(col_df$cols[which(col_df$study==dt_sub$study[1])], alpha.f = alpha1), bg = adjustcolor(col_df$cols[which(col_df$study==dt_sub$study[1])], alpha.f = alpha1)) # need bg for pch = 23
  }
}

# studies with randomly drawn sizes and survival
study_list1 <- study_list[which(study_list %in% unique(plot_dfS$study))]
for(i in 1:length(study_list1)){ # for each study
  study.i <- plot_dfS$study[which(plot_dfS$study==study_list1[i])][1]
  types.i <- unique(plot_dfS$data_type[which(plot_dfS$study==study.i)])
  for(j in 1:length(types.i)){ # for each data type
    types.ij <- types.i[j]
    dt_sub <- plot_dfS[which(plot_dfS$study==study.i & plot_dfS$data_type==types.i[j]),]
    jt <- rnorm(nrow(dt_sub), 0, jt_sd) # jitter the points
    lines(x = dt_sub$size_cm2_p, y = dt_sub$survived + jt, type = "p", pch = pch_fun(types.ij, "S"), col = adjustcolor(col_df$cols[which(col_df$study==dt_sub$study[1])], alpha.f = alpha1), bg = adjustcolor(col_df$cols[which(col_df$study==dt_sub$study[1])], alpha.f = alpha1)) # need bg for pch = 23
  }
}

# studies with randomly drawn survival
study_list1 <- study_list[which(study_list %in% unique(plot_dfIS$study))]
for(i in 1:length(study_list1)){ # for each study
  study.i <- plot_dfIS$study[which(plot_dfIS$study==study_list1[i])][1]
  types.i <- unique(plot_dfIS$data_type[which(plot_dfIS$study==study.i)])
  for(j in 1:length(types.i)){ # for each data type
    types.ij <- types.i[j]
    dt_sub <- plot_dfIS[which(plot_dfIS$study==study.i & plot_dfIS$data_type==types.i[j]),]
    jt <- rnorm(nrow(dt_sub), 0, jt_sd) # jitter the points
    lines(x = dt_sub$size_cm2_p, y = dt_sub$survived + jt, type = "p", pch = pch_fun(types.ij, "S"), col = adjustcolor(col_df$cols[which(col_df$study==dt_sub$study[1])], alpha.f = alpha1), bg = adjustcolor(col_df$cols[which(col_df$study==dt_sub$study[1])], alpha.f = alpha1)) # need bg for pch = 23
  }
}
legend(x = "topright", legend = c("individual data provided", "individual size and/or \nsurvival not provided"), pch = c(16, 1), col = "gray50", bty = "n",inset = c(0,0.15))
legend(x = "bottomright", legend = c("field", "nursery", "lab"), pch = c(f.pch, 18, l.pch), title = "Data type", col = "gray50", bg = "gray50", bty = "n", ncol = 2, inset = c(0,0.15))


# zoomed in 

plot(x = plot_dfI$size_cm2_p, y = plot_dfI$survived, type = "p", pch = 16, col = NA, las = 1, xlab = NA, ylab = NA, xlim  = c(0, 100))
mtext(side = 1, "Size (cm^2)", line = 2.5)
mtext(side = 2, "Prop. surviving", line = 2.5)
axis(side = 1, at = c(SC1, SC2, SC3, SC4, SC5), labels = NA, lwd.ticks = 2, col.ticks = "red")
# studies with individual data points
study_list1 <- study_list[which(study_list %in% unique(plot_dfI$study))]
for(i in 1:length(study_list1)){ # for each study
  study.i <- plot_dfI$study[which(plot_dfI$study==study_list1[i])][1]
  types.i <- unique(plot_dfI$data_type[which(plot_dfI$study==study.i)])
  for(j in 1:length(types.i)){ # for each data type
    types.ij <- types.i[j]
    dt_sub <- plot_dfI[which(plot_dfI$study==study.i & plot_dfI$data_type==types.i[j]),]
    jt <- rnorm(nrow(dt_sub), 0, jt_sd) # jitter the points
    lines(x = dt_sub$size_cm2_p, y = dt_sub$survived + jt, type = "p", pch = pch_fun(types.ij, "I"), col = adjustcolor(col_df$cols[which(col_df$study==dt_sub$study[1])], alpha.f = alpha1), bg = adjustcolor(col_df$cols[which(col_df$study==dt_sub$study[1])], alpha.f = alpha1)) # need bg for pch = 23
  }
}

# studies with randomly drawn sizes and survival
study_list1 <- study_list[which(study_list %in% unique(plot_dfS$study))]
for(i in 1:length(study_list1)){ # for each study
  study.i <- plot_dfS$study[which(plot_dfS$study==study_list1[i])][1]
  types.i <- unique(plot_dfS$data_type[which(plot_dfS$study==study.i)])
  for(j in 1:length(types.i)){ # for each data type
    types.ij <- types.i[j]
    dt_sub <- plot_dfS[which(plot_dfS$study==study.i & plot_dfS$data_type==types.i[j]),]
    jt <- rnorm(nrow(dt_sub), 0, jt_sd) # jitter the points
    lines(x = dt_sub$size_cm2_p, y = dt_sub$survived + jt, type = "p", pch = pch_fun(types.ij, "S"), col = adjustcolor(col_df$cols[which(col_df$study==dt_sub$study[1])], alpha.f = alpha1), bg = adjustcolor(col_df$cols[which(col_df$study==dt_sub$study[1])], alpha.f = alpha1)) # need bg for pch = 23
  }
}

# studies with randomly drawn survival
study_list1 <- study_list[which(study_list %in% unique(plot_dfIS$study))]
for(i in 1:length(study_list1)){ # for each study
  study.i <- plot_dfIS$study[which(plot_dfIS$study==study_list1[i])][1]
  types.i <- unique(plot_dfIS$data_type[which(plot_dfIS$study==study.i)])
  for(j in 1:length(types.i)){ # for each data type
    types.ij <- types.i[j]
    dt_sub <- plot_dfIS[which(plot_dfIS$study==study.i & plot_dfIS$data_type==types.i[j]),]
    jt <- rnorm(nrow(dt_sub), 0, jt_sd) # jitter the points
    lines(x = dt_sub$size_cm2_p, y = dt_sub$survived + jt, type = "p", pch = pch_fun(types.ij, "S"), col = adjustcolor(col_df$cols[which(col_df$study==dt_sub$study[1])], alpha.f = alpha1), bg = adjustcolor(col_df$cols[which(col_df$study==dt_sub$study[1])], alpha.f = alpha1)) # need bg for pch = 23
  }
}
legend(x = "topright", legend = c("individual data provided", "individual size and/or \nsurvival not provided"), pch = c(16, 1), col = "gray50", bty = "n",inset = c(0,0.15))
legend(x = "bottomright", legend = c("field", "nursery", "lab"), pch = c(f.pch, 18, l.pch), title = "Data type", col = "gray50", bg = "gray50", bty = "n", ncol = 2, inset = c(0,0.15))



# log scale

plot(x = log(plot_dfI$size_cm2_p + 1), y = plot_dfI$survived, type = "p", pch = 16, col = NA, las = 1, xlab = NA, ylab = NA, xaxt = "n")
axis(side = 1, at = c(0, log(c(10, 100, 10000))), labels = c(0, 10, 100, 10000))
axis(side = 1, at = log(c(SC1, SC2, SC3, SC4, SC5) + 1), labels = NA, lwd.ticks = 2, col.ticks = "red")
mtext(side = 1, "Size (cm^2)", line = 2.5)
mtext(side = 2, "Prop. surviving", line = 2.5)
# studies with individual data points
study_list1 <- study_list[which(study_list %in% unique(plot_dfI$study))]
for(i in 1:length(study_list1)){ # for each study
  study.i <- plot_dfI$study[which(plot_dfI$study==study_list1[i])][1]
  types.i <- unique(plot_dfI$data_type[which(plot_dfI$study==study.i)])
  for(j in 1:length(types.i)){ # for each data type
    types.ij <- types.i[j]
    dt_sub <- plot_dfI[which(plot_dfI$study==study.i & plot_dfI$data_type==types.i[j]),]
    jt <- rnorm(nrow(dt_sub), 0, jt_sd) # jitter the points
    lines(x = log(dt_sub$size_cm2_p + 1), y = dt_sub$survived + jt, type = "p", pch = pch_fun(types.ij, "I"), col = adjustcolor(col_df$cols[which(col_df$study==dt_sub$study[1])], alpha.f = alpha1), bg = adjustcolor(col_df$cols[which(col_df$study==dt_sub$study[1])], alpha.f = alpha1)) # need bg for pch = 23
  }
}

# studies with randomly drawn sizes and survival
study_list1 <- study_list[which(study_list %in% unique(plot_dfS$study))]
for(i in 1:length(study_list1)){ # for each study
  study.i <- plot_dfS$study[which(plot_dfS$study==study_list1[i])][1]
  types.i <- unique(plot_dfS$data_type[which(plot_dfS$study==study.i)])
  for(j in 1:length(types.i)){ # for each data type
    types.ij <- types.i[j]
    dt_sub <- plot_dfS[which(plot_dfS$study==study.i & plot_dfS$data_type==types.i[j]),]
    jt <- rnorm(nrow(dt_sub), 0, jt_sd) # jitter the points
    lines(x = log(dt_sub$size_cm2_p + 1), y = dt_sub$survived + jt, type = "p", pch = pch_fun(types.ij, "S"), col = adjustcolor(col_df$cols[which(col_df$study==dt_sub$study[1])], alpha.f = alpha1), bg = adjustcolor(col_df$cols[which(col_df$study==dt_sub$study[1])], alpha.f = alpha1)) # need bg for pch = 23
  }
}

# studies with randomly drawn survival
study_list1 <- study_list[which(study_list %in% unique(plot_dfIS$study))]
for(i in 1:length(study_list1)){ # for each study
  study.i <- plot_dfIS$study[which(plot_dfIS$study==study_list1[i])][1]
  types.i <- unique(plot_dfIS$data_type[which(plot_dfIS$study==study.i)])
  for(j in 1:length(types.i)){ # for each data type
    types.ij <- types.i[j]
    dt_sub <- plot_dfIS[which(plot_dfIS$study==study.i & plot_dfIS$data_type==types.i[j]),]
    jt <- rnorm(nrow(dt_sub), 0, jt_sd) # jitter the points
    lines(x = log(dt_sub$size_cm2_p + 1), y = dt_sub$survived + jt, type = "p", pch = pch_fun(types.ij, "S"), col = adjustcolor(col_df$cols[which(col_df$study==dt_sub$study[1])], alpha.f = alpha1), bg = adjustcolor(col_df$cols[which(col_df$study==dt_sub$study[1])], alpha.f = alpha1)) # need bg for pch = 23
  }
}

legend(x = "topright", legend = c("individual data provided", "individual size and/or \nsurvival not provided"), pch = c(16, 1), col = "gray50", bty = "n",inset = c(0,0.15))
legend(x = "bottomright", legend = c("field", "nursery", "lab"), pch = c(f.pch, 18, l.pch), title = "Data type", col = "gray50", bg = "gray50", bty = "n", ncol = 2, inset = c(0,0.15))


```


show with the proportion data as proportions instead


```{r}

plot_dfI <- apal_surv_ind1 %>% rename(size_cm2_p = size_cm2_1) %>% filter(is.na(disturbance)==T) # individual data
plot_dfS <- apal_surv_summ1 %>% mutate(size_cm2_p = if_else(is.na(size_cm2_mean)==F, size_cm2_mean, (size_cm2_min + size_cm2_max)/2))%>% filter(is.na(disturbance)==T)  # individual survival and size values drawn from summarized data
plot_dfIS <- test2[[1]] %>% rename(size_cm2_p = size_live_cm2)%>% filter(is.na(disturbance)==T)  # individual annual survival data drawn from summarized individual data

plot(x = log(plot_dfI$size_cm2_p + 1), y = plot_dfI$survived, type = "p", pch = 16, col = NA, las = 1, xlab = NA, ylab = NA, xaxt = "n")
axis(side = 1, at = c(0, log(c(10, 100, 10000))), labels = c(0, 10, 100, 10000))
axis(side = 1, at = log(c(SC1, SC2, SC3, SC4, SC5) + 1), labels = NA, lwd.ticks = 2, col.ticks = "red")
mtext(side = 1, "Size (cm^2)", line = 2.5)
mtext(side = 2, "Prop. surviving", line = 2.5)
# studies with individual data points
study_list1 <- study_list[which(study_list %in% unique(plot_dfI$study))]
for(i in 1:length(study_list1)){ # for each study
  study.i <- plot_dfI$study[which(plot_dfI$study==study_list1[i])][1]
  types.i <- unique(plot_dfI$data_type[which(plot_dfI$study==study.i)])
  for(j in 1:length(types.i)){ # for each data type
    types.ij <- types.i[j]
    dt_sub <- plot_dfI[which(plot_dfI$study==study.i & plot_dfI$data_type==types.i[j]),]
    jt <- rnorm(nrow(dt_sub), 0, jt_sd) # jitter the points
    lines(x = log(dt_sub$size_cm2_p + 1), y = dt_sub$survived + jt, type = "p", pch = pch_fun(types.ij, "I"), col = adjustcolor(col_df$cols[which(col_df$study==dt_sub$study[1])], alpha.f = alpha1), bg = adjustcolor(col_df$cols[which(col_df$study==dt_sub$study[1])], alpha.f = alpha1)) # need bg for pch = 23
  }
}

# studies with proportions
study_list1 <- study_list[which(study_list %in% unique(plot_dfS$study))]
for(i in 1:length(study_list1)){ # for each study
  study.i <- plot_dfS$study[which(plot_dfS$study==study_list1[i])][1]
  types.i <- unique(plot_dfS$data_type[which(plot_dfS$study==study.i)])
  for(j in 1:length(types.i)){ # for each data type
    types.ij <- types.i[j]
    dt_sub <- plot_dfS[which(plot_dfS$study==study.i & plot_dfS$data_type==types.i[j]),]
    lines(x = log(dt_sub$size_cm2_p + 1), y = dt_sub$prop_survived, type = "p", pch = pch_fun(types.ij, "S"), col = adjustcolor(col_df$cols[which(col_df$study==dt_sub$study[1])], alpha.f = 1), bg = adjustcolor(col_df$cols[which(col_df$study==dt_sub$study[1])], alpha.f = 1)) # need bg for pch = 23
  }
}

# studies with randomly drawn survival
study_list1 <- study_list[which(study_list %in% unique(plot_dfIS$study))]
for(i in 1:length(study_list1)){ # for each study
  study.i <- plot_dfIS$study[which(plot_dfIS$study==study_list1[i])][1]
  types.i <- unique(plot_dfIS$data_type[which(plot_dfIS$study==study.i)])
  for(j in 1:length(types.i)){ # for each data type
    types.ij <- types.i[j]
    dt_sub <- plot_dfIS[which(plot_dfIS$study==study.i & plot_dfIS$data_type==types.i[j]),]
    jt <- rnorm(nrow(dt_sub), 0, jt_sd) # jitter the points
    lines(x = log(dt_sub$size_cm2_p + 1), y = dt_sub$survived + jt, type = "p", pch = pch_fun(types.ij, "S"), col = adjustcolor(col_df$cols[which(col_df$study==dt_sub$study[1])], alpha.f = alpha1), bg = adjustcolor(col_df$cols[which(col_df$study==dt_sub$study[1])], alpha.f = alpha1)) # need bg for pch = 23
  }
}

legend(x = "topright", legend = c("individual data provided", "individual size and/or \nsurvival not provided"), pch = c(16, 1), col = "gray50", bty = "n",inset = c(0,0.2))
legend(x = "bottomright", legend = c("field", "nursery", "lab"), pch = c(f.pch, 18, l.pch), title = "Data type", col = "gray50", bg = "gray50", bty = "n", ncol = 2, inset = c(0,0.1))


```

```{r}
# apal_surv_summ1 %>% filter(study == "forrester_et_al_2013") %>% view()
# 
# 
# plot_dfS %>% filter(study == "forrester_et_al_2013") %>% view()

```


```{r}
# make a legend
col_df2 <- col_df %>% filter(study %in% unique(c(apal_surv_ind1$study, apal_surv_summ1$study))) # just the survival studies here

plot(x = rep(1, nrow(col_df2)), y = c(1:nrow(col_df2)), pch = 16, col = NA, xaxt = "n", yaxt = "n", ylab = NA, xlab = NA, bty = "n")
legend("right", legend = col_df2$study, col = col_df2$cols, pch = 15, bty = "n", ncol = 2)

```


fit binomial GLMM?


"brute force" resampling method

```{r}


# make a function for doing the resampling

# dtI = individual survival data
# dtS_list = list where each element is a replicate of survival data with random individual size and survival observations drawn from summarized values
# dtSI_list = list where each element is a replicate of survival data with random individual survival  observations drawn from summarized values (for data where individual sizes are known but survival at ~1 yr was estimated)

#dtS_list <- survi_fun(apal_surv_summ1[which(apal_surv_summ1$data_type=="field" & is.na(apal_surv_summ1$disturbance)==T),], n_reps = n_reps1, seed_n = 100)

#dtSI_list <- surv_indi_fun(apal_surv_ind_summ[which(apal_surv_ind_summ$data_type=="field" & is.na(apal_surv_ind_summ$disturbance)==T),], n_reps = n_reps1, seed_n = 200)

#dtI <- apal_surv_ind1 %>% filter(is.na(disturbance)==T, data_type == "field") # individual data

# n_reps = number of replicate datasets in dtS_list and dtIS_list
# n_sample = default number of observations to use in each subsample
resamp_surv3 <- function(dtI, dtS_list, dtIS_list, n_reps, n_sample, seed1){
  
  # check number of observations in each size class (repeat for each data set replicate for dtS_list; sizes won't change for dtIS_list since individual sizes are known and it's just the proportion surviving that varies)
  
  # dtI <- dtI %>% select(survived, size_class) %>% filter(is.na(size_class)==F)
  # 
  # # number of observations in each size class
  # dtI_n <- dtI %>% group_by(size_class) %>% summarize(n_SC = n())
  # # View(dtI_n)
  # 
  # dtIS_n <- dtIS_list[[1]] %>% group_by(size_class) %>% summarize(n_SC = n())
  # 
  #View(dtIS_n)
  
# get number of observations in each size class for each data replicate in the dtS_list
  
#   for(i in 1:n_reps){
#   
#   dtS_in <- dtS_list[[i]] %>% group_by(size_class) %>% summarize(n_SC = n()) %>% mutate(rep = i) 
#   
#   if(i == 1){
#     dtS_n <- dtS_in
#   } else{
#     
#     dtS_n <- rbind(dtS_n, dtS_in)
#   }
#   
# }
  
#View(dtS_n)

#View(dtS_list[[50]])
  
  # update: get the number of observations in each size class for the full data set (individual data + data from ith replicate of the summarized data)
  dtI <- dtI %>% select(survived, size_class) %>% filter(is.na(size_class)==F)
  
  #View(dtIS_i)
  
for(i in 1:n_reps){
  
  dtS_i <- dtS_list[[i]] %>% select(survived, size_class) %>% filter(is.na(size_class)==F)
  
  dtIS_i <- dtIS_list[[i]] %>% ungroup() %>% select(survived, size_class) %>% filter(is.na(size_class)==F)
  
  dt_in <- rbind(dtI, dtS_i, dtIS_i) %>% group_by(size_class) %>% summarize(n_SC = n()) %>% mutate(replicate = i)
  
  if(i == 1){
    dt_n <- dt_in
  } else{
    
    dt_n <- rbind(dt_n, dt_in)
  }
  
}
  
  #View(dt_n)
  
  # now calculate the number of observations to subsample from each size class for each iteration
  #n_sample <- min(n_sample, min(dtI_n$n_SC, dtIS_n$n_SC, dtS_n$n_SC))
  n_sample <- min(n_sample, min(dt_n$n_SC))
  
# for each data replicate iteration, get combined vector of survived or not from across the different data sets (vector of 0's and 1's) for each size class
# for each size class survival vector, get a vector of indexes, then randomly sample from these indexes. and keep track of which ones were sampled each iteration and remove them

# start with the individual data (only resampling, no iteration over dataset replicates)

  # SC_I <- list()
  # 
  # set.seed(seed1)
  # 
  # for(k in 1:5){ # for each size class
  #   
  #   # get the vectors of survival values
  #   S_SCk <- c(dtI$survived[which(dtI$size_class == k)]) 
  # 
  #   if(length(S_SCk) > 0){ # if this size class exists in the data
  #     
  #     # calculate the number of resampling iterations
  #     n_iter <- floor(dtI_n$n_SC[which(dtI_n$size_class==k)]/n_sample)  
  #     
  #     # get the index vector
  #     ind_x <- c(1:length(S_SCk))
  # 
  #     # holding vector for proportion surviving
  #      SC_v <- rep(NA, n_iter) # column = size class
  # 
  #     for(j in 1:n_iter){ # for each iteration
  #       
  #       # get random sample of observations
  #      sx <- sample(ind_x, size = n_sample, replace = F)
  # 
  # # calculate and store proportions surviving 
  # SC_v[j] <- length(which(S_SCk[sx] == 1))/n_sample
  # 
  # # update the vectors to avoid repeat sampling
  # ind_x <- ind_x[!(ind_x %in% sx)]
  # 
  #     }
  #     
  #   }
  #   
  #   SC_I[[k]] <- SC_v
  #   
  # }
  # 
  # 
  # repeat for the summarized data
  

  # update: include the individual data with the summarized data, otherwise the summarized data will be appearing n_rep times (with slightly different values each time) and the individual data will only appear once...alternative would be to keep the individual data separate and re-summarize the summarized data across the n_rep replicates and then combine?

set.seed(seed1)

# holding list
SC_surv <- list()

for(i in 1:n_reps){ # for each data replicate
  
  # holding list
SC_surv_i <- list()
SC_surv_i[[1]] <- NA
SC_surv_i[[2]] <- NA
SC_surv_i[[3]] <- NA
SC_surv_i[[4]] <- NA
SC_surv_i[[5]] <- NA
  
  # get the data set
  dtS_i <- dtS_list[[i]] %>% select(survived, size_class) %>% filter(is.na(size_class)==F)
  
  dtIS_i <- dtIS_list[[i]] %>% ungroup() %>% select(survived, size_class) %>% filter(is.na(size_class)==F)
  
  dt_i <- rbind(dtI, dtS_i, dtIS_i) 
  
  for(k in 1:5){ # for each size class
    
    # get the vector of survival values
  S_SCk <- c(dt_i$survived[which(dt_i$size_class == k)])  
    
    if(length(S_SCk) > 0){ # if this size class exists in the data
      
      n_iter <- floor(dt_n$n_SC[which(dt_n$size_class==k & dt_n$replicate==i)]/n_sample) # number of random replicates
      
      # get the index vector
     ind_x <- c(1:length(S_SCk))
     
      # holding vector for proportion surviving
       SC_v <- rep(NA, n_iter) # column = size class

      for(j in 1:n_iter){ # for each iteration

        # get random sample of observations
       sx <- sample(ind_x, size = n_sample, replace = F)

  # calculate and store proportions surviving
  SC_v[j] <- length(which(S_SCk[sx] == 1))/n_sample

  # update the vectors to avoid repeat sampling
  ind_x <- ind_x[!(ind_x %in% sx)]

      }

    SC_surv_i[[k]] <- SC_v
      
    }
    
  }
  
 
# store the list
SC_surv[[i]] <- SC_surv_i

  
} # end of iteration over data replicates


# want to make a data frame where columns are size class and proportion surviving
for(i in 1:n_reps){
  
  SC1_surv_df <- data.frame(
    prop_survived = SC_surv[[i]][[1]],
    size_class = 1,
    replicate = i,
    sample = c(1:length(SC_surv[[i]][[1]])),
    n = n_sample
  )
  
  SC2_surv_df <- data.frame(
    prop_survived = SC_surv[[i]][[2]],
    size_class = 2,
    replicate = i,
    sample = c(1:length(SC_surv[[i]][[2]])),
    n = n_sample
  )
  
  SC3_surv_df <- data.frame(
    prop_survived = SC_surv[[i]][[3]],
    size_class = 3,
    replicate = i,
    sample = c(1:length(SC_surv[[i]][[3]])),
    n = n_sample
  )
  
  SC4_surv_df <- data.frame(
    prop_survived = SC_surv[[i]][[4]],
    size_class = 4,
    replicate = i,
    sample = c(1:length(SC_surv[[i]][[4]])),
    n = n_sample
  )
  
  SC5_surv_df <- data.frame(
    prop_survived = SC_surv[[i]][[5]],
    size_class = 5,
    replicate = i,
    sample = c(1:length(SC_surv[[i]][[5]])),
    n = n_sample
  )
  
  SC_surv_dfi <- rbind(SC1_surv_df, SC2_surv_df, SC3_surv_df, SC4_surv_df, SC5_surv_df)

  if(i == 1){
    SC_surv_df <- SC_surv_dfi
  } else{
    SC_surv_df <- rbind(SC_surv_df, SC_surv_dfi)
  }
  
}


#View(SC_surv_df)

# convert to wide format

# also return summarized df with means and 95% confidence intervals for each size class with data
# columns = size class, mean, 5% and 95% quantiles

SC_surv_summ_df <- data.frame(
  size_class = c(1, 2, 3, 4, 5),
  mean = rep(NA, 5),
  Q05 = rep(NA, 5),
  Q95 = rep(NA, 5),
  tot_samples = rep(NA, 5), # total number of sampling events
  n_sample = n_sample # number of observations in each individual sampling event
)

for(k in 1:5){
  
  SC_surv_summ_df$mean[k] <- mean(SC_surv_df$prop_survived[which(SC_surv_df$size_class == k)], na.rm = T)
  SC_surv_summ_df$Q05[k] <- quantile(SC_surv_df$prop_survived[which(SC_surv_df$size_class == k)], 0.05, na.rm = T)
  SC_surv_summ_df$Q95[k] <- quantile(SC_surv_df$prop_survived[which(SC_surv_df$size_class == k)], 0.95, na.rm = T)
  SC_surv_summ_df$tot_samples[k] <- length(SC_surv_df$prop_survived[which(SC_surv_df$size_class == k)])
  
  
}

#View(SC_surv_summ_df)

return(list(SC_surv_df = SC_surv_df, SC_surv_summ_df = SC_surv_summ_df, n_sample = n_sample))
  
}




# repeat but for cases where there are only two data types (no dtIS)

resamp_surv2 <- function(dtI, dtS_list, n_reps, n_sample, seed1){
  
  # update: get the number of observations in each size class for the full data set (individual data + data from ith replicate of the summarized data)
  dtI <- dtI %>% select(survived, size_class) %>% filter(is.na(size_class)==F)
  
  #View(dtIS_i)
  
for(i in 1:n_reps){
  
  dtS_i <- dtS_list[[i]] %>% select(survived, size_class) %>% filter(is.na(size_class)==F)
  
  
  dt_in <- rbind(dtI, dtS_i) %>% group_by(size_class) %>% summarize(n_SC = n()) %>% mutate(replicate = i)
  
  if(i == 1){
    dt_n <- dt_in
  } else{
    
    dt_n <- rbind(dt_n, dt_in)
  }
  
}
  
  #View(dt_n)
  
  # now calculate the number of observations to subsample from each size class for each iteration
  n_sample <- min(n_sample, min(dt_n$n_SC))
  
# for each data replicate iteration, get combined vector of survived or not from across the different data sets (vector of 0's and 1's) for each size class
# for each size class survival vector, get a vector of indexes, then randomly sample from these indexes. and keep track of which ones were sampled each iteration and remove them



set.seed(seed1)

# holding list
SC_surv <- list()

for(i in 1:n_reps){ # for each data replicate
  
  # holding list
SC_surv_i <- list()
SC_surv_i[[1]] <- NA
SC_surv_i[[2]] <- NA
SC_surv_i[[3]] <- NA
SC_surv_i[[4]] <- NA
SC_surv_i[[5]] <- NA
  
  # get the data set
  dtS_i <- dtS_list[[i]] %>% select(survived, size_class) %>% filter(is.na(size_class)==F)
  
  dt_i <- rbind(dtI, dtS_i) 
  
  for(k in 1:5){ # for each size class
    
    # get the vector of survival values
  S_SCk <- c(dt_i$survived[which(dt_i$size_class == k)])  
    
    if(length(S_SCk) > 0){ # if this size class exists in the data
      
      n_iter <- floor(dt_n$n_SC[which(dt_n$size_class==k & dt_n$replicate==i)]/n_sample) # number of random replicates
      
      # get the index vector
     ind_x <- c(1:length(S_SCk))
     
      # holding vector for proportion surviving
       SC_v <- rep(NA, n_iter) # column = size class

      for(j in 1:n_iter){ # for each iteration

        # get random sample of observations
       sx <- sample(ind_x, size = n_sample, replace = F)

  # calculate and store proportions surviving
  SC_v[j] <- length(which(S_SCk[sx] == 1))/n_sample

  # update the vectors to avoid repeat sampling
  ind_x <- ind_x[!(ind_x %in% sx)]

      }

    SC_surv_i[[k]] <- SC_v
      
    }
    
  }
  
 
# store the list
SC_surv[[i]] <- SC_surv_i

  
} # end of iteration over data replicates


# want to make a data frame where columns are size class and proportion surviving
for(i in 1:n_reps){
  
  SC1_surv_df <- data.frame(
    prop_survived = SC_surv[[i]][[1]],
    size_class = 1,
    replicate = i,
    sample = c(1:length(SC_surv[[i]][[1]])),
    n = n_sample
  )
  
  SC2_surv_df <- data.frame(
    prop_survived = SC_surv[[i]][[2]],
    size_class = 2,
    replicate = i,
    sample = c(1:length(SC_surv[[i]][[2]])),
    n = n_sample
  )
  
  SC3_surv_df <- data.frame(
    prop_survived = SC_surv[[i]][[3]],
    size_class = 3,
    replicate = i,
    sample = c(1:length(SC_surv[[i]][[3]])),
    n = n_sample
  )
  
  SC4_surv_df <- data.frame(
    prop_survived = SC_surv[[i]][[4]],
    size_class = 4,
    replicate = i,
    sample = c(1:length(SC_surv[[i]][[4]])),
    n = n_sample
  )
  
  SC5_surv_df <- data.frame(
    prop_survived = SC_surv[[i]][[5]],
    size_class = 5,
    replicate = i,
    sample = c(1:length(SC_surv[[i]][[5]])),
    n = n_sample
  )
  
  SC_surv_dfi <- rbind(SC1_surv_df, SC2_surv_df, SC3_surv_df, SC4_surv_df, SC5_surv_df)

  if(i == 1){
    SC_surv_df <- SC_surv_dfi
  } else{
    SC_surv_df <- rbind(SC_surv_df, SC_surv_dfi)
  }
  
}


#View(SC_surv_df)

# also return summarized df with means and 95% confidence intervals for each size class with data
# columns = size class, mean, 5% and 95% quantiles

SC_surv_summ_df <- data.frame(
  size_class = c(1, 2, 3, 4, 5),
  mean = rep(NA, 5),
  Q05 = rep(NA, 5),
  Q95 = rep(NA, 5),
  tot_samples = rep(NA, 5), # total number of sampling events
  n_sample = n_sample # number of observations in each individual sampling event
)

for(k in 1:5){
  
  SC_surv_summ_df$mean[k] <- mean(SC_surv_df$prop_survived[which(SC_surv_df$size_class == k)], na.rm = T)
  SC_surv_summ_df$Q05[k] <- quantile(SC_surv_df$prop_survived[which(SC_surv_df$size_class == k)], 0.05, na.rm = T)
  SC_surv_summ_df$Q95[k] <- quantile(SC_surv_df$prop_survived[which(SC_surv_df$size_class == k)], 0.95, na.rm = T)
  SC_surv_summ_df$tot_samples[k] <- length(SC_surv_df$prop_survived[which(SC_surv_df$size_class == k)])
  
  
}

#View(SC_surv_summ_df)
  
return(list(SC_surv_df = SC_surv_df, SC_surv_summ_df = SC_surv_summ_df, n_sample = n_sample))
  
}






#x1 <- c(1, 2, 3, 4)

#x1[-1*c(1, 2)]




```



### field


```{r}
# get the random replicates
n_reps1 <- 50

surv_summ50_field <- survi_fun(apal_surv_summ1[which(apal_surv_summ1$data_type=="field" & is.na(apal_surv_summ1$disturbance)==T),], n_reps = n_reps1, seed_n = 100)

surv_ind_summ50_field <- surv_indi_fun(apal_surv_ind_summ[which(apal_surv_ind_summ$data_type=="field" & is.na(apal_surv_ind_summ$disturbance)==T),], n_reps = n_reps1, seed_n = 200)

surv_ind_field <- apal_surv_ind1 %>% filter(is.na(disturbance)==T, data_type == "field") # individual data


#View(surv_ind_summ50_field[[1]])


```


```{r}
# check where high size class 1 survival values are coming from
# View(apal_surv_summ1[which(apal_surv_summ1$size_class_mean == 1 & apal_surv_summ1$data_type=="field"),])
# 
# View(apal_surv_ind_summ[which(apal_surv_ind_summ$size_class == 1 & apal_surv_ind_summ$data_type=="field"),])
# 
# View(apal_surv_ind1[which(apal_surv_ind1$size_class == 1 & apal_surv_ind1$data_type=="field"),])


```



```{r}

n_sample1 <- 20

field_surv1 <- resamp_surv3(dtI = surv_ind_field, dtS_list = surv_summ50_field, dtIS_list = surv_ind_summ50_field, n_reps = n_reps1, n_sample = n_sample1, seed1 = 1000)


#View(field_surv1$SC_surv_df)

#View(field_surv1$SC_surv_summ_df)

#field_surv_wide <-field_surv1$SC_surv_df %>% pivot_wider(names_from = size_class, values_from = prop_survived)

#View(field_surv_wide)

#field_surv1$SC_surv_summ_df

```





```{r}

# test1 <- c(40, 30, 20, 80, 90, 60)
# 
# s_test <- sample(test1, size = 3, replace = F)
# 
# s_test
# 
# test1[!(test1 %in% s_test)]

```



```{r}
# plot(x = c(1:10), y = c(1:10))
# polygon(x = c(1, 2, 2, 1), y = c(5, 5, 7, 7))

```


```{r}
# plot results: plot just the field data, one data replicate as an example, put horizontal segments and confidence intervals showing the survival values

surv_df <- field_surv1$SC_surv_summ_df

plot_dfI <- surv_ind_field %>% rename(size_cm2_p = size_cm2_1) # individual data
plot_dfS <- surv_summ50_field[[1]] %>% rename(size_cm2_p = size_cm2)  # individual survival and size values drawn from summarized data
plot_dfIS <- surv_ind_summ50_field[[1]] %>% rename(size_cm2_p = size_live_cm2)  # individual annual survival data drawn from summarized individual data


plot(x = log(plot_dfI$size_cm2_p + 1), y = plot_dfI$survived, type = "p", pch = 16, col = NA, las = 1, xlab = NA, ylab = NA, xaxt = "n")
axis(side = 1, at = c(0, log(c(10, 100, 10000))), labels = c(0, 10, 100, 10000))
axis(side = 1, at = log(c(SC1, SC2, SC3, SC4, SC5) + 1), labels = NA, lwd.ticks = 2, col.ticks = "red")
mtext(side = 1, "Size (cm^2)", line = 2.5)
mtext(side = 2, "Prop. surviving", line = 2.5)

# studies with individual data points
study_list1 <- study_list[which(study_list %in% unique(plot_dfI$study))]
for(i in 1:length(study_list1)){ # for each study
  study.i <- plot_dfI$study[which(plot_dfI$study==study_list1[i])][1]
  types.i <- unique(plot_dfI$data_type[which(plot_dfI$study==study.i)])
  for(j in 1:length(types.i)){ # for each data type
    types.ij <- types.i[j]
    dt_sub <- plot_dfI[which(plot_dfI$study==study.i & plot_dfI$data_type==types.i[j]),]
    jt <- rnorm(nrow(dt_sub), 0, jt_sd) # jitter the points
    lines(x = log(dt_sub$size_cm2_p + 1), y = dt_sub$survived + jt, type = "p", pch = pch_fun(types.ij, "I"), col = adjustcolor(col_df$cols[which(col_df$study==dt_sub$study[1])], alpha.f = alpha1), bg = adjustcolor(col_df$cols[which(col_df$study==dt_sub$study[1])], alpha.f = alpha1)) # need bg for pch = 23
  }
}

# studies with randomly drawn sizes and survival
study_list1 <- study_list[which(study_list %in% unique(plot_dfS$study))]
for(i in 1:length(study_list1)){ # for each study
  study.i <- plot_dfS$study[which(plot_dfS$study==study_list1[i])][1]
  types.i <- unique(plot_dfS$data_type[which(plot_dfS$study==study.i)])
  for(j in 1:length(types.i)){ # for each data type
    types.ij <- types.i[j]
    dt_sub <- plot_dfS[which(plot_dfS$study==study.i & plot_dfS$data_type==types.i[j]),]
    jt <- rnorm(nrow(dt_sub), 0, jt_sd) # jitter the points
    lines(x = log(dt_sub$size_cm2_p + 1), y = dt_sub$survived + jt, type = "p", pch = pch_fun(types.ij, "S"), col = adjustcolor(col_df$cols[which(col_df$study==dt_sub$study[1])], alpha.f = alpha1), bg = adjustcolor(col_df$cols[which(col_df$study==dt_sub$study[1])], alpha.f = alpha1)) # need bg for pch = 23
  }
}

# studies with randomly drawn survival
study_list1 <- study_list[which(study_list %in% unique(plot_dfIS$study))]
for(i in 1:length(study_list1)){ # for each study
  study.i <- plot_dfIS$study[which(plot_dfIS$study==study_list1[i])][1]
  types.i <- unique(plot_dfIS$data_type[which(plot_dfIS$study==study.i)])
  for(j in 1:length(types.i)){ # for each data type
    types.ij <- types.i[j]
    dt_sub <- plot_dfIS[which(plot_dfIS$study==study.i & plot_dfIS$data_type==types.i[j]),]
    jt <- rnorm(nrow(dt_sub), 0, jt_sd) # jitter the points
    lines(x = log(dt_sub$size_cm2_p + 1), y = dt_sub$survived + jt, type = "p", pch = pch_fun(types.ij, "S"), col = adjustcolor(col_df$cols[which(col_df$study==dt_sub$study[1])], alpha.f = alpha1), bg = adjustcolor(col_df$cols[which(col_df$study==dt_sub$study[1])], alpha.f = alpha1)) # need bg for pch = 23
  }
}

# segments with survival estimates and polygons with confidence intervals
for(jj in 1:5){
  polygon(x = c(log(SC_lower[jj] + 1), log(SC_upper[jj] + 1), log(SC_upper[jj] + 1), log(SC_lower[jj] + 1)), y = c(rep(surv_df$Q05[which(surv_df$size_class==jj)], 2), rep(surv_df$Q95[which(surv_df$size_class==jj)], 2)), border = NA, col = adjustcolor("gray50", alpha.f = 0.25))
  segments(x0 = log(SC_lower[jj] + 1), y0 = surv_df$mean[which(surv_df$size_class==jj)], x1 = log(SC_upper[jj] + 1), lwd =1.5)
}


```


make legends

```{r}
# make a legend
col_df2 <- col_df %>% filter(study %in% unique(c(plot_dfI$study, plot_dfS$study))) # just the survival studies here

plot(x = rep(1, nrow(col_df2)), y = c(1:nrow(col_df2)), pch = 16, col = NA, xaxt = "n", yaxt = "n", ylab = NA, xlab = NA, bty = "n")
legend("right", legend = col_df2$study, col = col_df2$cols, pch = 15, bty = "n", ncol = 2)

plot(x = rep(1, nrow(col_df2)), y = c(1:nrow(col_df2)), pch = 16, col = NA, xaxt = "n", yaxt = "n", ylab = NA, xlab = NA, bty = "n")
legend(x = "topright", legend = c("individual data provided", "individual size and/or \nsurvival not provided"), pch = c(16, 1), col = "gray50", bty = "n",inset = c(0,0.2))


```

simplified version

```{r}

surv_df <- field_surv1$SC_surv_summ_df

plot_dfI <- apal_surv_ind1 %>% rename(size_cm2_p = size_cm2_1) %>% filter(is.na(disturbance)==T) # individual data
plot_dfS <- test1[[1]] %>% rename(size_cm2_p = size_cm2)%>% filter(is.na(disturbance)==T)  # individual survival and size values drawn from summarized data
plot_dfIS <- test2[[1]] %>% rename(size_cm2_p = size_live_cm2)%>% filter(is.na(disturbance)==T)  # individual annual survival data drawn from summarized individual data

plot(x = log(plot_dfI$size_cm2_p + 1), y = plot_dfI$survived, type = "p", pch = 16, col = NA, las = 1, xlab = NA, ylab = NA, xaxt = "n", yaxt = "n")
axis(side = 1, at = c(0, log(c(10, 100, 10000))), labels = c(0, 10, 100, 10000))
#axis(side = 2, at = c(0, 1), labels = c("Dead", "Alive"), las = 1)
axis(side = 2, at = c(0, 1), labels = c(0, 1), las = 1)
axis(side = 1, at = log(c(SC1, SC2, SC3, SC4, SC5) + 1), labels = NA, lwd.ticks = 3, col.ticks = "dodgerblue")
mtext(side = 1, expression("Size ("*cm^2*")"), line = 2.5)
mtext(side = 2, "Survival", line = 1)
# studies with individual data points
study_list1 <- study_list[which(study_list %in% unique(plot_dfI$study))]
for(i in 1:length(study_list1)){ # for each study
  study.i <- plot_dfI$study[which(plot_dfI$study==study_list1[i])][1]
  types.i <- unique(plot_dfI$data_type[which(plot_dfI$study==study.i)])
  for(j in 1:length(types.i)){ # for each data type
    types.ij <- types.i[j]
    dt_sub <- plot_dfI[which(plot_dfI$study==study.i & plot_dfI$data_type==types.i[j]),]
    jt <- rnorm(nrow(dt_sub), 0, jt_sd) # jitter the points
    lines(x = log(dt_sub$size_cm2_p + 1), y = dt_sub$survived + jt, type = "p", pch = 16, col = adjustcolor("black", alpha.f = 0.1)) # need bg for pch = 23
  }
}

# studies with randomly drawn sizes and survival
study_list1 <- study_list[which(study_list %in% unique(plot_dfS$study))]
for(i in 1:length(study_list1)){ # for each study
  study.i <- plot_dfS$study[which(plot_dfS$study==study_list1[i])][1]
  types.i <- unique(plot_dfS$data_type[which(plot_dfS$study==study.i)])
  for(j in 1:length(types.i)){ # for each data type
    types.ij <- types.i[j]
    dt_sub <- plot_dfS[which(plot_dfS$study==study.i & plot_dfS$data_type==types.i[j]),]
    jt <- rnorm(nrow(dt_sub), 0, jt_sd) # jitter the points
    lines(x = log(dt_sub$size_cm2_p + 1), y = dt_sub$survived + jt, type = "p", pch = 16, col = adjustcolor("black", alpha.f = 0.1)) # need bg for pch = 23
  }
}

# studies with randomly drawn survival
study_list1 <- study_list[which(study_list %in% unique(plot_dfIS$study))]
for(i in 1:length(study_list1)){ # for each study
  study.i <- plot_dfIS$study[which(plot_dfIS$study==study_list1[i])][1]
  types.i <- unique(plot_dfIS$data_type[which(plot_dfIS$study==study.i)])
  for(j in 1:length(types.i)){ # for each data type
    types.ij <- types.i[j]
    dt_sub <- plot_dfIS[which(plot_dfIS$study==study.i & plot_dfIS$data_type==types.i[j]),]
    jt <- rnorm(nrow(dt_sub), 0, jt_sd) # jitter the points
    lines(x = log(dt_sub$size_cm2_p + 1), y = dt_sub$survived + jt, type = "p", pch = 16, col = adjustcolor("black", alpha.f = 0.1)) # need bg for pch = 23
  }
}

# segments with survival estimates and polygons with confidence intervals
for(jj in 1:5){
  polygon(x = c(log(SC_lower[jj] + 1), log(SC_upper[jj] + 1), log(SC_upper[jj] + 1), log(SC_lower[jj] + 1)), y = c(rep(surv_df$Q05[which(surv_df$size_class==jj)], 2), rep(surv_df$Q95[which(surv_df$size_class==jj)], 2)), border = NA, col = adjustcolor("gray50", alpha.f = 0.25))
  segments(x0 = log(SC_lower[jj] + 1), y0 = surv_df$mean[which(surv_df$size_class==jj)], x1 = log(SC_upper[jj] + 1), lwd =1.5)
}

```



### nursery
repeat with nursery data

```{r}
# get the random replicates

surv_summ50_nurs <- survi_fun(apal_surv_summ1[which(apal_surv_summ1$data_type=="nursery" & is.na(apal_surv_summ1$disturbance)==T),], n_reps = n_reps1, seed_n = 100)

surv_ind_nurs <- apal_surv_ind1 %>% filter(is.na(disturbance)==T, data_type == "nursery") # individual data

nurs_surv1 <- resamp_surv2(dtI = surv_ind_nurs, dtS_list = surv_summ50_nurs, n_reps = n_reps1, n_sample = n_sample1, seed1 = 2000)

#View(nurs_surv1$SC_surv_summ_df) 
#nurs_surv1$n_sample
# size class 1 only has one 0 and since the sample number is 4 only 1 or 0.75 is possible. For size class 2, there are very few zeros and so since the sample number is 4, 0.75 should be the most common result after 1

```




```{r}

# plot results
surv_df <- nurs_surv1$SC_surv_summ_df

plot_dfI <- surv_ind_nurs %>% rename(size_cm2_p = size_cm2_1) # individual data
plot_dfS <- surv_summ50_nurs[[1]] %>% rename(size_cm2_p = size_cm2)  # individual survival and size values drawn from summarized data


plot(x = log(plot_dfI$size_cm2_p + 1), y = plot_dfI$survived, type = "p", pch = 16, col = NA, las = 1, xlab = NA, ylab = NA, xaxt = "n")
axis(side = 1, at = c(0, log(c(10, 100, 10000))), labels = c(0, 10, 100, 10000))
axis(side = 1, at = log(c(SC1, SC2, SC3, SC4, SC5) + 1), labels = NA, lwd.ticks = 2, col.ticks = "red")
mtext(side = 1, "Size (cm^2)", line = 2.5)
mtext(side = 2, "Prop. surviving", line = 2.5)

# studies with individual data points
study_list1 <- study_list[which(study_list %in% unique(plot_dfI$study))]
for(i in 1:length(study_list1)){ # for each study
  study.i <- plot_dfI$study[which(plot_dfI$study==study_list1[i])][1]
  types.i <- unique(plot_dfI$data_type[which(plot_dfI$study==study.i)])
  for(j in 1:length(types.i)){ # for each data type
    types.ij <- types.i[j]
    dt_sub <- plot_dfI[which(plot_dfI$study==study.i & plot_dfI$data_type==types.i[j]),]
    jt <- rnorm(nrow(dt_sub), 0, jt_sd) # jitter the points
    lines(x = log(dt_sub$size_cm2_p + 1), y = dt_sub$survived + jt, type = "p", pch = pch_fun(types.ij, "I"), col = adjustcolor(col_df$cols[which(col_df$study==dt_sub$study[1])], alpha.f = alpha1), bg = adjustcolor(col_df$cols[which(col_df$study==dt_sub$study[1])], alpha.f = alpha1)) # need bg for pch = 23
  }
}

# studies with randomly drawn sizes and survival
study_list1 <- study_list[which(study_list %in% unique(plot_dfS$study))]
for(i in 1:length(study_list1)){ # for each study
  study.i <- plot_dfS$study[which(plot_dfS$study==study_list1[i])][1]
  types.i <- unique(plot_dfS$data_type[which(plot_dfS$study==study.i)])
  for(j in 1:length(types.i)){ # for each data type
    types.ij <- types.i[j]
    dt_sub <- plot_dfS[which(plot_dfS$study==study.i & plot_dfS$data_type==types.i[j]),]
    jt <- rnorm(nrow(dt_sub), 0, jt_sd) # jitter the points
    lines(x = log(dt_sub$size_cm2_p + 1), y = dt_sub$survived + jt, type = "p", pch = pch_fun(types.ij, "S"), col = adjustcolor(col_df$cols[which(col_df$study==dt_sub$study[1])], alpha.f = alpha1), bg = adjustcolor(col_df$cols[which(col_df$study==dt_sub$study[1])], alpha.f = alpha1)) # need bg for pch = 23
  }
}


# segments with survival estimates and polygons with confidence intervals
for(jj in 1:5){
  polygon(x = c(log(SC_lower[jj] + 1), log(SC_upper[jj] + 1), log(SC_upper[jj] + 1), log(SC_lower[jj] + 1)), y = c(rep(surv_df$Q05[which(surv_df$size_class==jj)], 2), rep(surv_df$Q95[which(surv_df$size_class==jj)], 2)), border = NA, col = adjustcolor("gray50", alpha.f = 0.25))
  segments(x0 = log(SC_lower[jj] + 1), y0 = surv_df$mean[which(surv_df$size_class==jj)], x1 = log(SC_upper[jj] + 1), lwd =1.5)
}

```



```{r}
# make a legend
col_df2 <- col_df %>% filter(study %in% unique(c(plot_dfI$study, plot_dfS$study))) # just the survival studies here

plot(x = rep(1, nrow(col_df2)), y = c(1:nrow(col_df2)), pch = 16, col = NA, xaxt = "n", yaxt = "n", ylab = NA, xlab = NA, bty = "n")
legend("right", legend = col_df2$study, col = col_df2$cols, pch = 15, bty = "n", ncol = 1)

plot(x = rep(1, nrow(col_df2)), y = c(1:nrow(col_df2)), pch = 16, col = NA, xaxt = "n", yaxt = "n", ylab = NA, xlab = NA, bty = "n")
legend(x = "topright", legend = c("individual data provided", "individual size and/or \nsurvival not provided"), pch = c(n.pch, n.pchR), col = "gray50", pt.bg = "gray50", bty = "n",inset = c(0,0.2))


```

simplified version

```{r}

# plot results
surv_df <- nurs_surv1$SC_surv_summ_df

plot_dfI <- surv_ind_nurs %>% rename(size_cm2_p = size_cm2_1) # individual data
plot_dfS <- surv_summ50_nurs[[1]] %>% rename(size_cm2_p = size_cm2)  # individual survival and size values drawn from summarized data


plot(x = log(plot_dfI$size_cm2_p + 1), y = plot_dfI$survived, type = "p", pch = 16, col = NA, las = 1, xlab = NA, ylab = NA, xaxt = "n", xlim = c(0, log(100)), yaxt = "n")
axis(side = 1, at = c(0, log(c(10, 100, 10000))), labels = c(0, 10, 100, 10000))
axis(side = 1, at = log(c(SC1, SC2, SC3, SC4, SC5) + 1), labels = NA, lwd.ticks = 3, col.ticks = "dodgerblue")
mtext(side = 1, expression("Size ("*cm^2*")"), line = 2.5)
axis(side = 2, at = c(0, 1), labels = c(0, 1), las = 1)
mtext(side = 2, "Survival", line = 1)
# studies with individual data points
study_list1 <- study_list[which(study_list %in% unique(plot_dfI$study))]
for(i in 1:length(study_list1)){ # for each study
  study.i <- plot_dfI$study[which(plot_dfI$study==study_list1[i])][1]
  types.i <- unique(plot_dfI$data_type[which(plot_dfI$study==study.i)])
  for(j in 1:length(types.i)){ # for each data type
    types.ij <- types.i[j]
    dt_sub <- plot_dfI[which(plot_dfI$study==study.i & plot_dfI$data_type==types.i[j]),]
    jt <- rnorm(nrow(dt_sub), 0, jt_sd) # jitter the points
    lines(x = log(dt_sub$size_cm2_p + 1), y = dt_sub$survived + jt, type = "p", pch = 16, col = adjustcolor("black", alpha.f = 0.1)) # need bg for pch = 23
  }
}

# studies with randomly drawn sizes and survival
study_list1 <- study_list[which(study_list %in% unique(plot_dfS$study))]
for(i in 1:length(study_list1)){ # for each study
  study.i <- plot_dfS$study[which(plot_dfS$study==study_list1[i])][1]
  types.i <- unique(plot_dfS$data_type[which(plot_dfS$study==study.i)])
  for(j in 1:length(types.i)){ # for each data type
    types.ij <- types.i[j]
    dt_sub <- plot_dfS[which(plot_dfS$study==study.i & plot_dfS$data_type==types.i[j]),]
    jt <- rnorm(nrow(dt_sub), 0, jt_sd) # jitter the points
    lines(x = log(dt_sub$size_cm2_p + 1), y = dt_sub$survived + jt, type = "p",  pch = 16, col = adjustcolor("black", alpha.f = 0.1)) # need bg for pch = 23
  }
}


# segments with survival estimates and polygons with confidence intervals
for(jj in 1:5){
  polygon(x = c(log(SC_lower[jj] + 1), log(SC_upper[jj] + 1), log(SC_upper[jj] + 1), log(SC_lower[jj] + 1)), y = c(rep(surv_df$Q05[which(surv_df$size_class==jj)], 2), rep(surv_df$Q95[which(surv_df$size_class==jj)], 2)), border = NA, col = adjustcolor("gray50", alpha.f = 0.25))
  segments(x0 = log(SC_lower[jj] + 1), y0 = surv_df$mean[which(surv_df$size_class==jj)], x1 = log(SC_upper[jj] + 1), lwd =1.5)
}

```

```{r}
# turn the field and nursery data back into proportions and make bar plot comparing proportions surviving (with n = number of studies)

bp_FdfI <- surv_ind_field %>% select(size_class, survived) %>% filter(size_class < 3) # individual data # group_by(study, region, survey_yr, size_class)
bp_FdfS <- surv_summ50_field[[1]] %>% select(size_class, survived) %>% filter(size_class < 3) # individual survival and size values drawn from summarized data
bp_FdfIS <- surv_ind_summ50_field[[1]] %>% ungroup() %>% select(size_class, survived) %>% filter(size_class < 3) # individual annual survival data drawn from summarized individual data

#View(bp_FdfIS)

bp_Fdf <- rbind(bp_FdfI, bp_FdfS, bp_FdfIS) %>% group_by(size_class) %>% summarize(N_tot = n(), N_survived = sum(survived[survived == 1])) %>% mutate(prop_survived = N_survived/N_tot)
bp_Fdf$location <- "field"


#View(bp_FdfI)
#View(bp_Fdf)

bp_NdfI <- surv_ind_nurs %>% select(size_class, survived) # individual data
bp_NdfS <- surv_summ50_nurs[[1]] %>% select(size_class, survived) # individual data # individual survival and size values drawn from summarized data

#View(bp_NdfI)

bp_Ndf <- rbind(bp_NdfI, bp_NdfS) %>% group_by(size_class) %>% summarize(N_tot = n(), N_survived = sum(survived[survived == 1])) %>% mutate(prop_survived = N_survived/N_tot)
bp_Ndf$location <- "nursery"

bp_df <- rbind(bp_Fdf, bp_Ndf) %>% filter(is.na(size_class)==F)
bp_df$size_class <- ifelse(bp_df$size_class == 1, "0-10cm2", "10-100cm2")

```

```{r}
# bar plot

#View(bp_df)

bar.colors <- c("field" = "gray20", "nursery" = "gray80") 


ggplot(bp_df, aes(x=as.character(size_class), y=prop_survived, fill=location)) + 
   geom_bar(position = "dodge", stat = "identity") +
  xlab("Size class") +
  ylab("Proportion surviving") +
  scale_fill_manual(values=bar.colors)+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.text.x = element_text(size = 12), axis.text.y = element_text(size = 12), axis.title = element_text(size = 14), panel.border = element_rect(colour = "black", fill=NA, linewidth=0.8), plot.title = element_text(hjust=0, size = 14))
  #geom_boxplot()
  #geom_jitter() 

```


### lab
repeat with lab data

```{r}
# get the random replicates
surv_summ50_lab <- survi_fun(apal_surv_summ1[which(apal_surv_summ1$data_type=="lab" & is.na(apal_surv_summ1$disturbance)==T),], n_reps = n_reps1, seed_n = 100)

surv_ind_lab <- apal_surv_ind1 %>% filter(is.na(disturbance)==T, data_type == "lab") # individual data

lab_surv1 <- resamp_surv2(dtI = surv_ind_lab, dtS_list = surv_summ50_lab, n_reps = n_reps1, n_sample = n_sample1, seed1 = 4000)

lab_surv1$SC_surv_summ_df

```


plot results

```{r}

# plot results
surv_df <- lab_surv1$SC_surv_summ_df

plot_dfI <- surv_ind_lab %>% rename(size_cm2_p = size_cm2_1) # individual data
plot_dfS <- surv_summ50_lab[[1]] %>% rename(size_cm2_p = size_cm2)  # individual survival and size values drawn from summarized data


plot(x = log(plot_dfS$size_cm2_p + 1), y = plot_dfS$survived, type = "p", pch = 16, col = NA, las = 1, xlab = NA, ylab = NA, xaxt = "n", xlim = c(0, log(1 + 1)))
axis(side = 1, at = c(0, log(c(0.1, 1) + 1)), labels = c(0, exp(log(c(0.1, 1) + 1))-1))
axis(side = 1, at = log(c(SC1, SC2, SC3, SC4, SC5) + 1), labels = NA, lwd.ticks = 2, col.ticks = "red")
mtext(side = 1, "Size (cm^2)", line = 2.5)
mtext(side = 2, "Prop. surviving", line = 2.5)

# studies with individual data points
study_list1 <- study_list[which(study_list %in% unique(plot_dfI$study))]
for(i in 1:length(study_list1)){ # for each study
  study.i <- plot_dfI$study[which(plot_dfI$study==study_list1[i])][1]
  types.i <- unique(plot_dfI$data_type[which(plot_dfI$study==study.i)])
  for(j in 1:length(types.i)){ # for each data type
    types.ij <- types.i[j]
    dt_sub <- plot_dfI[which(plot_dfI$study==study.i & plot_dfI$data_type==types.i[j]),]
    jt <- rnorm(nrow(dt_sub), 0, jt_sd) # jitter the points
    lines(x = log(dt_sub$size_cm2_p + 1), y = dt_sub$survived + jt, type = "p", pch = pch_fun(types.ij, "I"), col = adjustcolor(col_df$cols[which(col_df$study==dt_sub$study[1])], alpha.f = alpha1), bg = adjustcolor(col_df$cols[which(col_df$study==dt_sub$study[1])], alpha.f = alpha1)) # need bg for pch = 23
  }
}

# studies with randomly drawn sizes and survival
study_list1 <- study_list[which(study_list %in% unique(plot_dfS$study))]
for(i in 1:length(study_list1)){ # for each study
  study.i <- plot_dfS$study[which(plot_dfS$study==study_list1[i])][1]
  types.i <- unique(plot_dfS$data_type[which(plot_dfS$study==study.i)])
  for(j in 1:length(types.i)){ # for each data type
    types.ij <- types.i[j]
    dt_sub <- plot_dfS[which(plot_dfS$study==study.i & plot_dfS$data_type==types.i[j]),]
    jt <- rnorm(nrow(dt_sub), 0, jt_sd) # jitter the points
    lines(x = log(dt_sub$size_cm2_p + 1), y = dt_sub$survived + jt, type = "p", pch = pch_fun(types.ij, "S"), col = adjustcolor(col_df$cols[which(col_df$study==dt_sub$study[1])], alpha.f = alpha1), bg = adjustcolor(col_df$cols[which(col_df$study==dt_sub$study[1])], alpha.f = alpha1)) # need bg for pch = 23
  }
}


# segments with survival estimates and polygons with confidence intervals
for(jj in 1:5){
  polygon(x = c(log(SC_lower[jj] + 1), log(SC_upper[jj] + 1), log(SC_upper[jj] + 1), log(SC_lower[jj] + 1)), y = c(rep(surv_df$Q05[which(surv_df$size_class==jj)], 2), rep(surv_df$Q95[which(surv_df$size_class==jj)], 2)), border = NA, col = adjustcolor("gray50", alpha.f = 0.25))
  segments(x0 = log(SC_lower[jj] + 1), y0 = surv_df$mean[which(surv_df$size_class==jj)], x1 = log(SC_upper[jj] + 1), lwd =1.5)
}

```

```{r}
# make a legend
col_df2 <- col_df %>% filter(study %in% unique(c(plot_dfI$study, plot_dfS$study))) # just the survival studies here

plot(x = rep(1, nrow(col_df2)), y = c(1:nrow(col_df2)), pch = 16, col = NA, xaxt = "n", yaxt = "n", ylab = NA, xlab = NA, bty = "n")
legend("right", legend = col_df2$study, col = col_df2$cols, pch = 15, bty = "n", ncol = 1)

plot(x = rep(1, nrow(col_df2)), y = c(1:nrow(col_df2)), pch = 16, col = NA, xaxt = "n", yaxt = "n", ylab = NA, xlab = NA, bty = "n")
legend(x = "topright", legend = c("individual data provided", "individual size and/or \nsurvival not provided"), pch = c(l.pch, l.pchR), col = "gray50", pt.bg = "gray50", bty = "n",inset = c(0,0.2))


```

### export

```{r}

saveRDS(field_surv1, "parameter_lists/field_surv_pars.rds")
saveRDS(nurs_surv1, "parameter_lists/nurs_surv_pars.rds")
saveRDS(lab_surv1, "parameter_lists/lab_surv_pars.rds")

```



## growth

growth: as a function of initial size, calculate how high your growth rate would need to be to grow into the next size class, two size classes up, etc. and how low it would need to be to shrink to each smaller size class, then calculate proportions of observations that are above this line (transition), below the shrinkage line (shrink) or inbetween (stay in same size class)

```{r}

# function for calculating threshold growth/shrinkage rates need to reach a neighboring size class
# s0 = initial coral size
# SC = lower threshold of a larger size class (for growth) or upper threshold of a smaller size class (for shrinkage)
GS_thresh <- function(s0, SC){
  
  if(s0 > SC){ # if looking at shrinkage
    
    rate_min <- s0-SC # min shrinkage rate needed to shrink into this size class in one year
  }
  
  if(s0 < SC){ # if looking at growth
    rate_min <- SC-s0
  }
  
  if(s0 == SC){
    rate_min <- 0
  }
  
  return(rate_min)
}

GS_thresh <- Vectorize(GS_thresh, "s0")

```



```{r}

s0_1 <- seq(from = SC_lower[1], to = SC_upper[1], length.out = 100) # starting in first size class
s0_2 <- seq(from = SC_lower[2], to = SC_upper[2], length.out = 100) # second class
s0_3 <- seq(from = SC_lower[3], to = SC_upper[3], length.out = 100) # third class
s0_4 <- seq(from = SC_lower[4], to = SC_upper[4], length.out = 100) # fourth class
s0_5 <- seq(from = SC_lower[5], to = SC_upper[5], length.out = 100) # fifth class

plot(x = s0_1, y = GS_thresh(s0_1, SC_lower[2]), type = "l", las = 1)


```


```{r}
# make a function whose inputs are initial size and growth rate and whose outputs are the size class the individual will be in 1 yr later

# or really all you need to do is add the annual growth rate to the current size to get the size one year later, and then make a column with the size class in the next year (and remember to check for negative sizes and assign them as zero)

```


```{r}
# function for estimating individual growth from the summarized data without information on individual sizes


growthi_fun <- function(summ_growth_df, n_reps, seed_n){
  
  set.seed(seed_n) # set the seed
  
  data_list <- list() # holding list for the data sets
  
  for(i in 1:n_reps){ # for each realization
    
   for(j in 1:nrow(summ_growth_df)){
     
     # number of individuals that went into the jth observation of proportion surviving
     n_ind <- summ_growth_df$n_final[j]
     
     # make the holding data set to fill in with individual values (and keep all the metadata)
     data.ij <- summ_growth_df %>% select(study, region, location, plot, treatment_1, treatment_2, latitude, longitude, depth_m, survey_yr, data_type, size_live, fragment, time_interval_yr, disturbance)
     data.ij <- data.ij[j,] # want just the jth row
     
     # replicate all the metadata n_ind times
     data.ij <- data.ij[rep(seq_len(nrow(data.ij)), n_ind), ]
     
     # now fill in size and individual growth columns
     
     # holding column for size:
     data.ij$size_cm2 <- rep(NA, nrow(data.ij))
     
     # holding column for growth
     data.ij$growth_cm2_yr <- rep(NA, nrow(data.ij))
     
     # fill in sizes by drawing a random size for each individual
      # if there is no information on variability (size_sd = NA)
     if(is.na(summ_growth_df$size_cm2_sd[j])==T){
       
       # if only mean size is known, just draw individual survival rates and assume they all have the same size
       if(is.na(summ_growth_df$size_cm2_min[j])==T){
         data.ij$size_cm2 <- rep(summ_growth_df$size_cm2_mean[j], n_ind)
         
       } else{ # if only range is known or if only mean/midpoint and range is known, draw from uniform distribution
         data.ij$size_cm2 <- runif(n_ind, min = summ_growth_df$size_cm2_min[j], max = summ_growth_df$size_cm2_max[j])
       }
      
       
     } else{ # if there is information on size variability
       
       if(is.na(summ_growth_df$size_cm2_min[j])==F){ # if mean size, sd, and range are known, draw sizes from a normal distribution and adjust to make sure they are within the given range
         
         x1 <- which(10 == 1) # start with length 0
         
         while(length(x1) < n_ind){
           # draw sizes from a normal distribution
           x1 <- c(x1, rnorm(n_ind-length(x1), summ_growth_df$size_cm2_mean[j], summ_growth_df$size_cm2_sd[j]))
           # make sure these values all fall within the observed size range:
           x1 <- x1[which(x1>= summ_growth_df$size_cm2_min[j] & x1 <= summ_growth_df$size_cm2_max[j])]
         }
         
         data.ij$size_cm2 <- x1
         
       } else{ # if mean size and sd are known but there is no information on range, draw sizes from a normal distribution and adjust to make sure they are non-zero
         
         x1 <- which(10 == 1) # start with length 0
         
         while(length(x1) < n_ind){
           # draw sizes from a normal distribution
           x1 <- c(x1, rnorm(n_ind-length(x1), summ_growth_df$size_cm2_mean[j], summ_growth_df$size_cm2_sd[j]))
           # make sure these values all fall within the observed size range:
           x1 <- x1[which(x1>= 0)]
         }
         
         data.ij$size_cm2 <- x1
         
       }
     
       
     }
       
     
     # now calculate growth rates
     
     # if there is no information on variability (growth_sd = NA)
     if(is.na(summ_growth_df$growth_cm2_yr_sd[j])==T){
       
       # if only mean growth is known, just assume all the individuals have the same growth rate
       if(is.na(summ_growth_df$growth_cm2_yr_min[j])==T){
         data.ij$growth_cm2_yr <- rep(summ_growth_df$growth_cm2_yr_mean[j], n_ind)
         
       } else{ # if only range is known or if only mean/midpoint and range is known, draw from uniform distribution
         data.ij$growth_cm2_yr <- runif(n_ind, min = summ_growth_df$growth_cm2_yr[j], max = summ_growth_df$growth_cm2_yr[j])
       }
      
       
     } else{ # if there is information on size variability
       
       if(is.na(summ_growth_df$growth_cm2_yr_min[j])==F){ # if mean size, sd, and range are known, draw sizes from a normal distribution and adjust to make sure they are within the given range
         
         x1 <- which(10 == 1) # start with length 0
         
         while(length(x1) < n_ind){
           # draw sizes from a normal distribution
           x1 <- c(x1, rnorm(n_ind-length(x1), summ_growth_df$growth_cm2_yr_mean[j], summ_growth_df$growth_cm2_yr_sd[j]))
           # make sure these values all fall within the observed size range:
           x1 <- x1[which(x1>= summ_growth_df$growth_cm2_yr_min[j] & x1 <= summ_growth_df$growth_cm2_yr_max[j])]
         }
         
         data.ij$growth_cm2_yr <- x1
         
       } else{ # if mean size and sd are known but there is no information on range, draw sizes from a normal distribution (and it's ok if they are less than 0 since that would just mean shrinkage)
         
         data.ij$growth_cm2_yr <- rnorm(n_ind, summ_growth_df$growth_cm2_yr_mean[j], summ_growth_df$growth_cm2_yr_sd[j])
         
         
       }
     
       
     }
     
     if(j == 1){
       data.j <- data.ij
     } else{
       
       data.j <- rbind(data.j, data.ij)
       
     }
    
     
   } # end of iteration over rows in the data frame
    
    # store this data frame in the list
    data_list[[i]] <- data.j
    
  } # end of iteration over random replicates
  
  
  # return the list of data frames
  
  return(data_list)
}

#View(apal_growth_summ)

```


```{r}

dtS_in <- apal_growth_summ1 %>% filter(data_type == "field" & is.na(disturbance)==T)
dtS_list <- growthi_fun(dtS_in, n_reps = 50, seed_n = 500)

#View(dtS_list[[1]])
dtI <- apal_growth_ind2 %>% filter(data_type == "field" & is.na(disturbance)==T)

#View(dtI)

dtS_in <- apal_growth_summ1 %>% filter(data_type == "nursery" & is.na(disturbance)==T)
dtS_list <- growthi_fun(dtS_in, n_reps = 50, seed_n = 500)

#View(dtS_list[[1]])
dtI <- apal_growth_ind2 %>% filter(data_type == "nursery" & is.na(disturbance)==T)


```



```{r}

# function for resampling to get transition matrix parameter estimates

resamp_growth2 <- function(dtI, dtS_list, n_reps, n_sample, seed1){
  
  
  # prep individual data
dtI <- dtI %>% mutate(size_class = SC_funV(size_cm2_1)) %>% mutate(size_cm2_t1 = size_cm2_1 + growth_cm2_yr_1) %>% mutate(size_cm2_t1 = if_else(size_cm2_t1 < 0, 0, size_cm2_t1)) %>%  mutate(SC_t1 = SC_funV(size_cm2_t1)) %>% select(size_class, SC_t1) # individual data
  
  # get number of observations in each size class (repeat for each data set replicate)
  
# individual data
dtI_n <- dtI %>% select(size_class) %>% filter(is.na(size_class)==F)
  
  for(i in 1:n_reps){
  
    # ith realization from summarized data
  dtS_i <- dtS_list[[i]] %>% mutate(size_class = SC_funV(size_cm2)) %>% select(size_class) %>% filter(is.na(size_class)==F)
  
  
  dt_in <- rbind(dtI_n, dtS_i) %>% group_by(size_class) %>% summarize(n_SC = n()) %>% mutate(replicate = i)
  
  if(i == 1){
    dt_n <- dt_in
  } else{
    
    dt_n <- rbind(dt_n, dt_in)
  }
  
}
  
  #View(dt_n)
  
  # now calculate the number of observations to subsample from each size class for each iteration
  n_sample <- min(n_sample, min(dt_n$n_SC))

# for each data replicate iteration, get combined vector of survived or not from across the different data sets (vector of 0's and 1's) for each size class
# for each size class survival vector, get a vector of indexes, then randomly sample from these indexes. and keep track of which ones were sampled each iteration and remove them

# note the individual data are going to be re-used each data replicate... but there's still variation since the exact observations being sampled will vary

# outer holding list for the transition matrices
T_mat_all <- list()

# outer holding list for the number of resampling iterations
n_iter_all <- list()

set.seed(seed1) # set the seed

for(i in 1:n_reps){ # for each data replicate
  
#n_iter <- floor(min_mat[i,1]/n_sample) # number of random replicates 


# prep summarized growth data for this data replicate
 dtS_list_i <- dtS_list[[i]] %>% mutate(size_class = SC_funV(size_cm2)) %>% mutate(size_cm2_t1 = size_cm2 + growth_cm2_yr) %>% mutate(size_cm2_t1 = if_else(size_cm2_t1 < 0, 0, size_cm2_t1)) %>%  mutate(SC_t1 = SC_funV(size_cm2_t1)) %>% select(size_class, SC_t1)
  
 # combine with the individual data to make a single data set
 dt_i <- rbind(dtI, dtS_list_i)
 
 n_iter <- rep(NA, 5) # holding vector for number of resampling iterations
 S_SCX <- rep(NA, 5) # holding vector for number of observations in each size class
 
 
 for(k in 1:5){ # for each size class
   
    # get the number of individuals starting in this size class
  S_SCX[k] <- length(which(dt_i$size_class==k))
    
  # get the number of random replicate iterations
  if(S_SCX[k] > 0){
    n_iter[k] <- floor(dt_n$n_SC[which(dt_n$size_class==k & dt_n$replicate==i)]/n_sample)
  } else{
    n_iter[k] <- 0
  }
  
 }
 
 #View(dt_i)
 

# determine which size classes have lengths > 0
SC_present <- which(S_SCX !=0)

# hodling list for index vectors
ind_list <- list()

for(k in 1:5){
  
  ind_list[[k]] <- ifelse(rep(S_SCX[k], S_SCX[k])!=0, c(1:S_SCX[k]), NA) # indexes for kth size class are 1:number of individuals in that size class

}

 T_mat_list <- list() # holding list for the transition matrices from each subsample

for(j in 1:max(n_iter)){ # for each sampling iteration
  
  # holding matrix for the transition matrix
  T_mat <- matrix(NA, nrow = 5, ncol = 5) # column = initial size class, row = final size class
  
  for(k in SC_present){ # for each size class present
    
    
    if(j <= n_iter[k]){
      
    # get random sample for each size class 
  sx <- sample(ind_list[[k]], size = n_sample, replace = F)
  
  dt_k <- dt_i %>% filter(size_class == k)
  
  dt_k <- dt_k[sx,]
  
  flag_NA <- 0
  
    } else{
      
      flag_NA <- 1
    }

  # total that went to size class 1
    SC_tot1 <- ifelse(flag_NA == 0, length(dt_k$SC_t1[which(dt_k$SC_t1==1)]), NA)
    
    # total that went to size class 2
    SC_tot2 <- ifelse(flag_NA == 0, length(dt_k$SC_t1[which(dt_k$SC_t1==2)]), NA)
    
    # total that went to size class 3
    SC_tot3 <- ifelse(flag_NA == 0, length(dt_k$SC_t1[which(dt_k$SC_t1==3)]), NA)
    
    # total that went to size class 4
    SC_tot4 <- ifelse(flag_NA == 0, length(dt_k$SC_t1[which(dt_k$SC_t1==4)]), NA)
    
    # total that went to size class 5
    SC_tot5 <- ifelse(flag_NA == 0, length(dt_k$SC_t1[which(dt_k$SC_t1==5)]), NA)
    
    SC_props <- c(SC_tot1, SC_tot2, SC_tot3, SC_tot4, SC_tot5)/n_sample
  
    T_mat[ ,k] <- SC_props # store transition probabilities for kth size class
    
  # update the vectors to avoid repeat sampling
  ind_list[[k]] <- ind_list[[k]][!(ind_list[[k]] %in% sx)]
  
  
  } # end of iteration over each size class for jth data subset

  T_mat_list[[j]] <- T_mat
  
} # end of iteration over each data subset


T_mat_all[[i]] <- T_mat_list
n_iter_all[[i]] <- n_iter

} # end of iteration over each dataset replicate

# want to make make a summarized data frame with mean and quantiles for each value
# first get each them in matrices
#n_iter_max <- floor(max(min_mat[,1])/n_sample) # max number of random replicates

mat1 <- matrix(NA, nrow = 1, ncol = 5) # rows = replicates, cols = from 1 to 1, 1 to 2, 1 to 3, etc.
mat2 <- matrix(NA, nrow = 1, ncol = 5) # from 2 to 1, 2 to 2, 2 to 3, etc.
mat3 <- matrix(NA, nrow = 1, ncol = 5) # from 3 to 1, 2, 3, 4, 5
mat4 <- matrix(NA, nrow = 1, ncol = 5) # from 4 to 1, 2, 3, 4, 5
mat5 <- matrix(NA, nrow = 1, ncol = 5) # from 5 to 1, 2, 3, 4, 5

mat_list <- list(mat1, mat2, mat3, mat4, mat5)

matn <- matrix(NA, nrow = 50, ncol = 5) # number of resampling iterations for size class 1, 2, 3, 4, 5

for(i in 1:n_reps){
  n_iter <- n_iter_all[[i]] # number of resampling iterations for each size class

  matn[i,] <- n_iter
  
  for(k in SC_present){
    
    for(j in 1:n_iter[k]){
    
    if(i == 1 & j == 1){
      
      mat_list[[k]] <- T_mat_all[[i]][[j]][,k]
      
    } else{
      
      # rbind
      mat_list[[k]] <- rbind(mat_list[[k]], T_mat_all[[i]][[j]][,k])
      
    } # end of if else
    
  } # end of j loop
   
  } # end of k loop
  
} # end of i loop

#mat1

# then calculate the mean and quantiles of each column in each matrix


summ_list <- list()

for(k in SC_present){
  
summ_k <- data.frame(
  mean = rep(NA, 5),
  Q05 = rep(NA, 5),
  Q95 = rep(NA, 5)
)

summ_k$mean <- apply(mat_list[[k]], MARGIN = 2, FUN = mean) # mean probabilities of transitioning to size class 1, 2, 3, 4, 5. NOTE they don't necessarily sum to 1 so can't just plug them into the model
summ_k$Q05 <- apply(mat_list[[k]], MARGIN = 2, FUN = quantile, probs = 0.05)
summ_k$Q95 <- apply(mat_list[[k]], MARGIN = 2, FUN = quantile, probs = 0.95)
  
summ_list[[k]] <- summ_k

}



return(list(T_mat_all = T_mat_all, mat_list = mat_list, matn = matn, summ_list = summ_list, n_sample = n_sample)) #T_mat_summ = T_mat_summ, 
  
}




```


```{r}
T_mat_all[[1]][[1]][,1]

#as.vector(T_mat_all[[1]][[1]])

```


apal_growth_summ1

apal_growth_ind2



```{r}
dtS_in <- apal_growth_summ1 %>% filter(data_type == "field" & is.na(disturbance)==T)


dtS_list <- growthi_fun(dtS_in, n_reps = 50, seed_n = 500)

#View(dtS_list[[1]])

dtI <- apal_growth_ind2 %>% filter(data_type == "field" & is.na(disturbance)==T)


test1 <- resamp_growth2(dtI, dtS_list, n_reps = n_reps1, n_sample = n_sample1, seed1)


#test1$matn

#test1$mat_list[[1]]

test1$summ_list[[4]]

```



### field

field data only:


```{r}
# get the random replicates and the growth estimates
n_reps1 <- 50

dtS_in <- apal_growth_summ1 %>% filter(data_type == "field" & is.na(disturbance)==T)

growth_summ50_field <- growthi_fun(dtS_in, n_reps = n_reps1, seed_n = 500)

growth_ind_field <- apal_growth_ind2 %>% filter(data_type == "field" & is.na(disturbance)==T)# individual data


n_sample1 <- 20

field_growth1 <- resamp_growth2(dtI = growth_ind_field, dtS_list = growth_summ50_field, n_reps = n_reps1, n_sample = n_sample1, seed1 = 200)



```





```{r}
# plot the field growth data

plot_dfI <- growth_ind_field %>% mutate(size_cm2_t1 = size_cm2 + growth_cm2_yr) %>% mutate(size_cm2_t1 = if_else(size_cm2_t1 < 0, 0, size_cm2_t1)) %>% rename(size_cm2_p = size_cm2_1)# individual data

plot_dfS <- growth_summ50_field[[1]] %>%  mutate(size_cm2_t1 = size_cm2 + growth_cm2_yr) %>% mutate(size_cm2_t1 = if_else(size_cm2_t1 < 0, 0, size_cm2_t1)) %>% rename(size_cm2_p = size_cm2) # individual survival and size values drawn from summarized data

par(pty = "s")
plot(x = log(plot_dfI$size_cm2_p + 1), y = log(plot_dfI$size_cm2_t1 + 1), type = "p", pch = 16, col = NA, las = 1, xlab = NA, ylab = NA, xaxt = "n", yaxt = "n", asp = 1) #, xaxs = "i", yaxs = "i"
axis(side = 1, at = c(0, log(c(10, 100, 10000))), labels = c(0, 10, 100, 10000))
axis(side = 1, at = log(c(SC1, SC2, SC3, SC4, SC5) + 1), labels = NA, lwd.ticks = 2, col.ticks = "red")
axis(side = 2, at = c(0, log(c(10, 100, 10000))), labels = c(0, 10, 100, 10000), las = 1)
axis(side = 2, at = log(c(SC1, SC2, SC3, SC4, SC5) + 1), labels = NA, lwd.ticks = 2, col.ticks = "red")
abline(a = 0, b = 1)
abline(v = log(SC_lower[2:5] + 1), col = "red", lty = 2)
abline(h = log(SC_lower[2:5] + 1), col = "red", lty = 2)

mtext(side = 1, "Size at time t (cm^2)", line = 2.5)
mtext(side = 2, "Size one year later", line = 3.25)

# studies with individual data points
study_list1 <- study_list[which(study_list %in% unique(plot_dfI$study))]
for(i in 1:length(study_list1)){ # for each study
  study.i <- plot_dfI$study[which(plot_dfI$study==study_list1[i])][1]
  types.i <- unique(plot_dfI$data_type[which(plot_dfI$study==study.i)])
  for(j in 1:length(types.i)){ # for each data type
    types.ij <- types.i[j]
    dt_sub <- plot_dfI[which(plot_dfI$study==study.i & plot_dfI$data_type==types.i[j]),]
    jt <- rnorm(nrow(dt_sub), 0, jt_sd) # jitter the points
    lines(x = log(dt_sub$size_cm2_p + 1), y = log(dt_sub$size_cm2_t1 + 1), type = "p", pch = pch_fun(types.ij, "I"), col = adjustcolor(col_df$cols[which(col_df$study==dt_sub$study[1])], alpha.f = alpha1), bg = adjustcolor(col_df$cols[which(col_df$study==dt_sub$study[1])], alpha.f = alpha1)) # need bg for pch = 23
  }
}

# studies with randomly drawn sizes and survival
study_list1 <- study_list[which(study_list %in% unique(plot_dfS$study))]
for(i in 1:length(study_list1)){ # for each study
  study.i <- plot_dfS$study[which(plot_dfS$study==study_list1[i])][1]
  types.i <- unique(plot_dfS$data_type[which(plot_dfS$study==study.i)])
  for(j in 1:length(types.i)){ # for each data type
    types.ij <- types.i[j]
    dt_sub <- plot_dfS[which(plot_dfS$study==study.i & plot_dfS$data_type==types.i[j]),]
    jt <- rnorm(nrow(dt_sub), 0, jt_sd) # jitter the points
    lines(x = log(dt_sub$size_cm2_p + 1), y = log(dt_sub$size_cm2_t1 + 1), type = "p", pch = pch_fun(types.ij, "S"), col = adjustcolor(col_df$cols[which(col_df$study==dt_sub$study[1])], alpha.f = alpha1), bg = adjustcolor(col_df$cols[which(col_df$study==dt_sub$study[1])], alpha.f = alpha1)) # need bg for pch = 23
  }
}


```


check the zeros

```{r}
#plot_dfI$coral_id[which(plot_dfI$size_cm2_t1 == 0)] # check on these -- might be an issue with extrapolating growth rate to a year?

#noaa_dt4 %>% filter(TagID=="CF3-t973") %>% view()

#noaa_dt4 %>% filter(TagID=="GR3-t398") %>% view()

# I think the zeros are cases where something extreme happened (e.g., branches broke off, large amount of tissue mortality) and the time interval was less than one year so when that shrinkage rate got extrapolated to a year it caused the predicted size to go below zero

```


```{r}
# make a legend
col_df2 <- col_df %>% filter(study %in% unique(c(plot_dfI$study, plot_dfS$study))) # just the survival studies here

plot(x = rep(1, nrow(col_df2)), y = c(1:nrow(col_df2)), pch = 16, col = NA, xaxt = "n", yaxt = "n", ylab = NA, xlab = NA, bty = "n")
legend("right", legend = col_df2$study, col = col_df2$cols, pch = 15, bty = "n", ncol = 1)

plot(x = rep(1, nrow(col_df2)), y = c(1:nrow(col_df2)), pch = 16, col = NA, xaxt = "n", yaxt = "n", ylab = NA, xlab = NA, bty = "n")
legend(x = "topright", legend = c("individual data provided", "individual size and/or \ngrowth not provided"), pch = c(f.pch, f.pchR), col = "gray50", pt.bg = "gray50", bty = "n",inset = c(0,0.2))


```

### nursery
nursery data only:

```{r}
# get the random replicates and the growth estimates
n_reps1 <- 50

dtS_in <- apal_growth_summ1 %>% filter(data_type == "nursery" & is.na(disturbance)==T)

growth_summ50_nurs <- growthi_fun(dtS_in, n_reps = n_reps1, seed_n = 500)

growth_ind_nurs <- apal_growth_ind2 %>% filter(data_type == "nursery" & is.na(disturbance)==T)# individual data


n_sample1 <- 20

nurs_growth1 <- resamp_growth2(dtI = growth_ind_nurs, dtS_list = growth_summ50_nurs, n_reps = n_reps1, n_sample = n_sample1, seed1 = 200)



```




```{r}
# plot the nursery growth data

plot_dfI <- growth_ind_nurs %>% mutate(size_cm2_t1 = size_cm2 + growth_cm2_yr) %>% mutate(size_cm2_t1 = if_else(size_cm2_t1 < 0, 0, size_cm2_t1)) %>% rename(size_cm2_p = size_cm2_1)# individual data

#plot_dfI %>% filter(size_cm2_t1 == 0) %>% view() # check on these -- might be an issue with extrapolating growth rate to a year?

plot_dfS <- growth_summ50_nurs[[1]] %>% mutate(size_cm2_t1 = size_cm2 + growth_cm2_yr) %>% mutate(size_cm2_t1 = if_else(size_cm2_t1 < 0, 0, size_cm2_t1))%>% rename(size_cm2_p = size_cm2) # individual survival and size values drawn from summarized data

par(pty = "s")
plot(x = log(plot_dfI$size_cm2_p + 1), y = log(plot_dfI$size_cm2_t1 + 1), type = "p", pch = 16, col = NA, las = 1, xlab = NA, ylab = NA, xaxt = "n", yaxt = "n", asp = 1, ylim = c(0, max(c(log(plot_dfI$size_cm2_t1 + 1), log(plot_dfS$size_cm2_t1 + 1)))), xlim = c(0, max(c(log(plot_dfI$size_cm2_t1 + 1), log(plot_dfS$size_cm2_t1 + 1))))) #, xaxs = "i", yaxs = "i"
axis(side = 1, at = c(0, log(c(10, 100, 10000))), labels = c(0, 10, 100, 10000))
axis(side = 1, at = log(c(SC1, SC2, SC3, SC4, SC5) + 1), labels = NA, lwd.ticks = 2, col.ticks = "red")
axis(side = 2, at = c(0, log(c(10, 100, 10000))), labels = c(0, 10, 100, 10000), las = 1)
axis(side = 2, at = log(c(SC1, SC2, SC3, SC4, SC5) + 1), labels = NA, lwd.ticks = 2, col.ticks = "red")
abline(a = 0, b = 1)
abline(v = log(SC_lower[2:5] + 1), col = "red", lty = 2)
abline(h = log(SC_lower[2:5] + 1), col = "red", lty = 2)

mtext(side = 1, "Size at time t (cm^2)", line = 2.5)
mtext(side = 2, "Size one year later", line = 3.25)

# studies with individual data points
study_list1 <- study_list[which(study_list %in% unique(plot_dfI$study))]
for(i in 1:length(study_list1)){ # for each study
  study.i <- plot_dfI$study[which(plot_dfI$study==study_list1[i])][1]
  types.i <- unique(plot_dfI$data_type[which(plot_dfI$study==study.i)])
  for(j in 1:length(types.i)){ # for each data type
    types.ij <- types.i[j]
    dt_sub <- plot_dfI[which(plot_dfI$study==study.i & plot_dfI$data_type==types.i[j]),]
    jt <- rnorm(nrow(dt_sub), 0, jt_sd) # jitter the points
    lines(x = log(dt_sub$size_cm2_p + 1), y = log(dt_sub$size_cm2_t1 + 1), type = "p", pch = pch_fun(types.ij, "I"), col = adjustcolor(col_df$cols[which(col_df$study==dt_sub$study[1])], alpha.f = alpha1), bg = adjustcolor(col_df$cols[which(col_df$study==dt_sub$study[1])], alpha.f = alpha1)) # need bg for pch = 23
  }
}

# studies with randomly drawn sizes and survival
study_list1 <- study_list[which(study_list %in% unique(plot_dfS$study))]
for(i in 1:length(study_list1)){ # for each study
  study.i <- plot_dfS$study[which(plot_dfS$study==study_list1[i])][1]
  types.i <- unique(plot_dfS$data_type[which(plot_dfS$study==study.i)])
  for(j in 1:length(types.i)){ # for each data type
    types.ij <- types.i[j]
    dt_sub <- plot_dfS[which(plot_dfS$study==study.i & plot_dfS$data_type==types.i[j]),]
    jt <- rnorm(nrow(dt_sub), 0, jt_sd) # jitter the points
    lines(x = log(dt_sub$size_cm2_p + 1), y = log(dt_sub$size_cm2_t1 + 1), type = "p", pch = pch_fun(types.ij, "S"), col = adjustcolor(col_df$cols[which(col_df$study==dt_sub$study[1])], alpha.f = alpha1), bg = adjustcolor(col_df$cols[which(col_df$study==dt_sub$study[1])], alpha.f = alpha1)) # need bg for pch = 23
  }
}


```


```{r}
# make a legend
col_df2 <- col_df %>% filter(study %in% unique(c(plot_dfI$study, plot_dfS$study))) # just the survival studies here

plot(x = rep(1, nrow(col_df2)), y = c(1:nrow(col_df2)), pch = 16, col = NA, xaxt = "n", yaxt = "n", ylab = NA, xlab = NA, bty = "n")
legend("right", legend = col_df2$study, col = col_df2$cols, pch = 15, bty = "n", ncol = 1)

plot(x = rep(1, nrow(col_df2)), y = c(1:nrow(col_df2)), pch = 16, col = NA, xaxt = "n", yaxt = "n", ylab = NA, xlab = NA, bty = "n")
legend(x = "topright", legend = c("individual data provided", "individual size and/or \ngrowth not provided"), pch = c(n.pch, n.pchR), col = "gray50", pt.bg = "gray50", bty = "n",inset = c(0,0.2))


```

### lab

lab data only

```{r}
# get the random replicates and the growth estimates
n_reps1 <- 50

dtS_in <- apal_growth_summ1 %>% filter(data_type == "lab" & is.na(disturbance)==T)

growth_summ50_lab <- growthi_fun(dtS_in, n_reps = n_reps1, seed_n = 500)

growth_ind_lab <- apal_growth_ind2 %>% filter(data_type == "lab" & is.na(disturbance)==T)# individual data


n_sample1 <- 20

lab_growth1 <- resamp_growth2(dtI = growth_ind_lab, dtS_list = growth_summ50_lab, n_reps = n_reps1, n_sample = n_sample1, seed1 = 200)



```



```{r}
# plot the lab growth data

plot_dfI <- growth_ind_lab %>% mutate(size_cm2_t1 = size_cm2 + growth_cm2_yr) %>% mutate(size_cm2_t1 = if_else(size_cm2_t1 < 0, 0, size_cm2_t1)) %>% rename(size_cm2_p = size_cm2_1)# individual data

#plot_dfI %>% filter(size_cm2_t1 == 0) %>% view() # check on these -- might be an issue with extrapolating growth rate to a year?

plot_dfS <- growth_summ50_lab[[1]] %>% mutate(size_cm2_t1 = size_cm2 + growth_cm2_yr) %>% mutate(size_cm2_t1 = if_else(size_cm2_t1 < 0, 0, size_cm2_t1))%>% rename(size_cm2_p = size_cm2) # individual survival and size values drawn from summarized data

par(pty = "s")
plot(x = log(plot_dfI$size_cm2_p + 1), y = log(plot_dfI$size_cm2_t1 + 1), type = "p", pch = 16, col = NA, las = 1, xlab = NA, ylab = NA, xaxt = "n", yaxt = "n", asp = 1, ylim = c(0, max(c(log(plot_dfI$size_cm2_t1 + 1), log(plot_dfS$size_cm2_t1 + 1)))), xlim = c(0, max(c(log(plot_dfI$size_cm2_t1 + 1), log(plot_dfS$size_cm2_t1 + 1))))) #, xaxs = "i", yaxs = "i"
axis(side = 1, at = c(0, log(c(10, 100, 10000))), labels = c(0, 10, 100, 10000))
axis(side = 1, at = log(c(SC1, SC2, SC3, SC4, SC5) + 1), labels = NA, lwd.ticks = 2, col.ticks = "red")
axis(side = 2, at = c(0, log(c(10, 100, 10000))), labels = c(0, 10, 100, 10000), las = 1)
axis(side = 2, at = log(c(SC1, SC2, SC3, SC4, SC5) + 1), labels = NA, lwd.ticks = 2, col.ticks = "red")
abline(a = 0, b = 1)
abline(v = log(SC_lower[2:5] + 1), col = "red", lty = 2)
abline(h = log(SC_lower[2:5] + 1), col = "red", lty = 2)

mtext(side = 1, "Size at time t (cm^2)", line = 2.5)
mtext(side = 2, "Size one year later", line = 3.25)

# studies with individual data points
study_list1 <- study_list[which(study_list %in% unique(plot_dfI$study))]
for(i in 1:length(study_list1)){ # for each study
  study.i <- plot_dfI$study[which(plot_dfI$study==study_list1[i])][1]
  types.i <- unique(plot_dfI$data_type[which(plot_dfI$study==study.i)])
  for(j in 1:length(types.i)){ # for each data type
    types.ij <- types.i[j]
    dt_sub <- plot_dfI[which(plot_dfI$study==study.i & plot_dfI$data_type==types.i[j]),]
    jt <- rnorm(nrow(dt_sub), 0, jt_sd) # jitter the points
    lines(x = log(dt_sub$size_cm2_p + 1), y = log(dt_sub$size_cm2_t1 + 1), type = "p", pch = pch_fun(types.ij, "I"), col = adjustcolor(col_df$cols[which(col_df$study==dt_sub$study[1])], alpha.f = alpha1), bg = adjustcolor(col_df$cols[which(col_df$study==dt_sub$study[1])], alpha.f = alpha1)) # need bg for pch = 23
  }
}

# studies with randomly drawn sizes and survival
study_list1 <- study_list[which(study_list %in% unique(plot_dfS$study))]
for(i in 1:length(study_list1)){ # for each study
  study.i <- plot_dfS$study[which(plot_dfS$study==study_list1[i])][1]
  types.i <- unique(plot_dfS$data_type[which(plot_dfS$study==study.i)])
  for(j in 1:length(types.i)){ # for each data type
    types.ij <- types.i[j]
    dt_sub <- plot_dfS[which(plot_dfS$study==study.i & plot_dfS$data_type==types.i[j]),]
    jt <- rnorm(nrow(dt_sub), 0, jt_sd) # jitter the points
    lines(x = log(dt_sub$size_cm2_p + 1), y = log(dt_sub$size_cm2_t1 + 1), type = "p", pch = pch_fun(types.ij, "S"), col = adjustcolor(col_df$cols[which(col_df$study==dt_sub$study[1])], alpha.f = alpha1), bg = adjustcolor(col_df$cols[which(col_df$study==dt_sub$study[1])], alpha.f = alpha1)) # need bg for pch = 23
  }
}


```

```{r}
# make a legend
col_df2 <- col_df %>% filter(study %in% unique(c(plot_dfI$study, plot_dfS$study))) # just the survival studies here

plot(x = rep(1, nrow(col_df2)), y = c(1:nrow(col_df2)), pch = 16, col = NA, xaxt = "n", yaxt = "n", ylab = NA, xlab = NA, bty = "n")
legend("right", legend = col_df2$study, col = col_df2$cols, pch = 15, bty = "n", ncol = 1)

plot(x = rep(1, nrow(col_df2)), y = c(1:nrow(col_df2)), pch = 16, col = NA, xaxt = "n", yaxt = "n", ylab = NA, xlab = NA, bty = "n")
legend(x = "topright", legend = c("individual data provided", "individual size and/or \ngrowth not provided"), pch = c(l.pch, l.pchR), col = "gray50", pt.bg = "gray50", bty = "n",inset = c(0,0.2))


```





### export

```{r}

saveRDS(field_growth1, "parameter_lists/field_growth_pars.rds")
saveRDS(nurs_growth1, "parameter_lists/nurs_growth_pars.rds")
saveRDS(lab_growth1, "parameter_lists/lab_growth_pars.rds")

```




















